% ═══════════════════════════════════════════════════════════════════════════════
% G_F BVP ALL-GATES PASS + PHYSICAL PRIORS BOX
% ═══════════════════════════════════════════════════════════════════════════════
%
% File: gf_bvp_allgates_physical_priors_box.tex
% Created: 2026-01-29
% Issue: OPR-21b/c — BVP pipeline results with physical priors
% Status: Framework GREEN; values YELLOW (BVP-gated, tuned priors)
%
% Usage: \input{../../edc_papers/_shared/boxes/gf_bvp_allgates_physical_priors_box}
%
% Dependencies: tcolorbox, booktabs, enumitem
%
% ═══════════════════════════════════════════════════════════════════════════════

\begin{tcolorbox}[breakable, enhanced, colback=green!5!white, colframe=green!60!black,
    title={\textbf{BVP Pipeline: All Gates Pass (OPR-21b/c)}}, width=\linewidth]

\textbf{Result:} Three independent gates PASS simultaneously at parameters found via scan.
This match is \textbf{YELLOW} until priors ($\delta$, $d_{\text{LR}}$, fw) are derived from 5D action.

\vspace{0.5em}
\textbf{Why non-trivial:} At baseline parameters, Gate~1 failed by factor $38\times$.
After parameter scan, $X_{\text{EDC}}/X_{\text{target}} = 1.045$ --- a $36.8\times$ improvement.
Gate~1 ($I_4$ overlap) is the \textbf{primary falsification channel}: if mode-overlap mechanism
cannot produce the required suppression, the entire $G_F$ derivation path fails.

\vspace{0.5em}
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{llcc}
\toprule
\textbf{Gate} & \textbf{Test} & \textbf{Value} & \textbf{Status} \\
\midrule
Gate 1 & $I_4$ overlap window & $2.08 \times 10^{-3}$ GeV & \textcolor{OliveGreen}{\textbf{PASS}} (was $38\times$ off) \\
Gate 2 & $M_{\text{eff}}$ mass scaling & $2.43$ GeV & \textcolor{OliveGreen}{\textbf{PASS}} \\
Gate 3 & $g_{\text{eff}}^2$ coupling & $0.196$ & \textcolor{OliveGreen}{\textbf{PASS}} \\
\midrule
\multicolumn{2}{l}{\textbf{Target match:}} & \multicolumn{2}{l}{$X_{\text{EDC}} / X_{\text{target}} = 1.045$ (\textbf{4.5\% error})} \\
\bottomrule
\end{tabular}
\end{center}

\vspace{0.5em}
\textbf{Tuned parameters} \tagDc{}+\tagCal{}:
\begin{center}
\begin{tabular}{lccl}
\toprule
\textbf{Parameter} & \textbf{Value} & \textbf{Physical Length} & \textbf{Interpretation} \\
\midrule
$\delta$ (brane thickness) & $0.533\,\text{GeV}^{-1}$ & $0.105\,\text{fm}$ & $= \hbar/(2m_p)$ \tagDc{} \\
LR\_separation & $8.0\,\delta$ & $0.84\,\text{fm}$ & $\approx r_p$ (suggestive) \\
fermion\_width & $0.8\,\delta$ & $0.085\,\text{fm}$ & $\approx 0.4\,\lambda_N$ \\
\bottomrule
\end{tabular}
\end{center}

\vspace{0.5em}
\textbf{Sensitivity decomposition} \tagDer{}:
\begin{itemize}[nosep]
    \item LR\_separation elasticity: $-6.5$ (dominant, exponential control)
    \item fermion\_width elasticity: $+1.3$ (secondary, polynomial control)
\end{itemize}
Physical interpretation: LR controls the exponential overlap suppression; fw fine-tunes the residual.

\vspace{0.5em}
\textbf{Coincidence flag:} $d_{\text{LR}} = 8\delta = 0.84\,\text{fm} \approx r_p = 0.84\,\text{fm}$.\\
\textit{Marked as suggestive, NOT derived. Would require deriving $d_{\text{LR}}$ from chiral localization.}

\vspace{0.5em}
\textbf{Potential-shape robustness:} Gaussian, RS-like, and tanh backgrounds all produce wells with depth $\sim 1/\delta^2$ and width $\sim \delta$. Specific shape is modeling choice \tagDc{}, not uniquely determined.

\end{tcolorbox}

% ═══════════════════════════════════════════════════════════════════════════════
% EPISTEMIC GUARDRAIL
% ═══════════════════════════════════════════════════════════════════════════════

\begin{tcolorbox}[colback=yellow!10!white, colframe=yellow!70!black,
    title={\textbf{Epistemic Status: Framework GREEN, Values YELLOW}}]

\textbf{What this validates:}
\begin{itemize}[nosep]
    \item Mode-overlap mechanism can produce the required $G_F$ suppression \checkmark
    \item Three independent consistency gates are satisfied simultaneously \checkmark
    \item Parameter space exists where BVP solution matches target \checkmark
    \item Sensitivity is understood: LR dominant (exponential), fw secondary \checkmark
\end{itemize}

\textbf{What this does NOT validate:}
\begin{itemize}[nosep]
    \item Physical priors are not derived from 5D action (tuned) $\times$
    \item The $d_{\text{LR}} \approx r_p$ coincidence is not explained $\times$
    \item Potential shape is ansatz, not derived from bulk dynamics $\times$
    \item $\delta = \hbar/(2m_p)$ is postulated, not derived $\times$
\end{itemize}

\vspace{0.5em}
\textbf{Big Guardrail:}
\begin{center}
\fbox{\parbox{0.9\linewidth}{\centering
\textbf{Framework: GREEN} (mechanism works)\\
\textbf{Values: YELLOW} (BVP-gated, tuned priors --- not yet derived)
}}
\end{center}

\vspace{0.5em}
\textbf{Upgrade path YELLOW $\to$ GREEN:}
\begin{enumerate}[nosep]
    \item Derive $\delta = \hbar/(2m_p)$ from 5D action (no tuning)
    \item Derive $d_{\text{LR}}$ from chiral localization / junction geometry
    \item Derive fw from BVP stability or eigenvalue constraints
\end{enumerate}

\textbf{Cross-references:}
\begin{itemize}[nosep]
    \item Sensitivity analysis: \texttt{docs/GF\_BVP\_TUNING\_DECOMPOSITION.md}
    \item Physical priors: \texttt{docs/GF\_BVP\_PHYSICAL\_PRIORS.md}
    \item Gate report: \texttt{docs/GF\_BVP\_GATE\_REPORT.md}
    \item Defense notes: \texttt{docs/GF\_BVP\_DEFENSE\_NOTES.md}
\end{itemize}

\end{tcolorbox}

% ═══════════════════════════════════════════════════════════════════════════════
% END OF BOX
% ═══════════════════════════════════════════════════════════════════════════════
