% ==============================================================================
% Chapter 6: Neutrinos as Edge Modes
% Status: [Dc]/[P] — Mass suppression identified, mixing postulated
% ==============================================================================

\section{Neutrinos as Edge Modes}
\label{sec:ch6_neutrinos}

\begin{tcolorbox}[edcGuardrail, title=\textbf{Epistemic Status}]
This chapter explains neutrino properties within EDC's 5D framework:
\begin{itemize}[nosep]
    \item Neutrino smallness from edge-mode overlap suppression \tagDc{}
    \item Three flavors from $\mathbb{Z}_3 \subset \mathbb{Z}_6$ mode structure \tagI{}
    \item PMNS mixing from generation wavefunction overlaps \tagP{}
\end{itemize}
\textbf{What is NOT claimed:} Explicit mass values are not derived. PMNS angles
are postulated, not computed. Mass hierarchy origin remains (open).
\end{tcolorbox}

% ==============================================================================
% FRAMEWORK 2.0 LANGUAGE COMPLIANCE
% ==============================================================================
\begin{tcolorbox}[colback=blue!3!white, colframe=blue!50!black,
    title=\textbf{Framework 2.0 Language Compliance}]
\small
\textbf{EDC Projection Principle:} Every physical process has a \textbf{5D bulk+brane cause}
whose observable residue is a \textbf{3D shadow} on the observer boundary.

\textbf{In this chapter:}
\begin{itemize}[nosep]
    \item \textbf{5D cause:} Neutrino as edge mode at brane–bulk interface.
    \item \textbf{Brane process:} Suppressed overlap with Higgs profile in brane interior.
    \item \textbf{3D shadow:} Tiny neutrino mass ($m_\nu/m_e \sim 10^{-6}$); left-chirality.
\end{itemize}

\textbf{Standard Model} treats neutrino smallness as a see-saw mechanism input;
EDC proposes it as a geometric consequence of boundary localization.
\end{tcolorbox}

% ------------------------------------------------------------------------------
% PHYSICAL PROCESS NARRATIVE (Feynman-style)
% ------------------------------------------------------------------------------

\begin{tcolorbox}[colback=green!5!white, colframe=green!50!black,
    title=\textbf{Physical Process Narrative: How Neutrinos Become ``Ghostly''}]
\textbf{What physically happens, step by step:}

\textbf{Step 1: The brane has a boundary.}
In EDC, our observable 3D universe sits at the ``observer face'' of a thick membrane.
This membrane has finite thickness $\delta$ along the fifth dimension $\xi$. At $\xi = 0$,
there is an \emph{interface}---a boundary separating the brane interior (where electrons
and quarks live) from the 5D bulk (the Plenum). This interface is where neutrinos
reside \tagP{}.

\textbf{Step 2: Edge modes are boundary-trapped excitations.}
In condensed matter physics, edge modes are states that exist \emph{only} at boundaries---they
cannot propagate into the bulk. In EDC, the neutrino is proposed to be such a mode:
its wavefunction $\psi_\nu(\xi)$ peaks sharply at the interface ($\xi \approx 0$) and
decays exponentially into the brane interior \tagP{}.

\textbf{Step 3: The Higgs lives in the interior, not at the edge.}
The mass-generating mechanism (Higgs field profile $h(\xi)$) is localized in the brane
interior ($\xi > 0$), not at the interface. Electrons and quarks, which live in the
interior, have large overlap with $h(\xi)$ and acquire ``normal'' masses. Neutrinos,
trapped at the interface, have exponentially \emph{suppressed} overlap \tagDc{}.

\textbf{Step 4: Suppressed overlap $\Rightarrow$ tiny mass.}
The effective 4D mass comes from the overlap integral $m \sim \int |\psi(\xi)|^2 h(\xi) d\xi$.
For the electron (interior mode), this integral is $\sim 1$. For the neutrino (edge mode),
it is $\sim e^{-\Delta z/\kappa^{-1}}$ where $\Delta z$ is the separation and $\kappa^{-1}$
is the penetration depth. With $\Delta z/\kappa^{-1} \approx 14$, we get $m_\nu/m_e \sim 10^{-6}$
\tagDc{}.

\textbf{Step 5: Same boundary conditions $\Rightarrow$ same chirality.}
Chapter~\ref{ch:va_structure} showed that the asymmetric mass profile at the brane
boundary selects left-handed fermions. The neutrino, being an edge mode at this same
boundary, inherits the same chirality selection. This is why weak interactions are
V$-$A: only $\nu_L$ couples \tagDc{}.

\textbf{Step 6: Three flavors from three angular sectors.}
The $\mathbb{Z}_3$ structure (Ch.~\ref{ch:three_generations}) provides three inequivalent
angular positions around the hexagonal axis. Each neutrino flavor occupies one sector:
$\nu_e$ at $\phi = 0$, $\nu_\mu$ at $\phi = 2\pi/3$, $\nu_\tau$ at $\phi = 4\pi/3$ \tagI{}.

\textbf{Step 7: Mixing = angular overlap.}
The PMNS matrix elements are overlap integrals between flavor wavefunctions (different
angular sectors) and mass eigenstates. Large angles (e.g., $\theta_{23} \approx 45°$)
indicate significant angular overlap; small angles (e.g., $\theta_{13} \approx 8.5°$)
indicate well-separated sectors \tagP{}.

\textbf{What would falsify this?}
\begin{itemize}[nosep]
    \item Large neutrino mass ($m_\nu > 1$ eV): would require $\Delta z/\kappa^{-1} < 14$,
          inconsistent with electron mass explanation
    \item Right-handed weak coupling: boundary conditions would be wrong
    \item 4th light neutrino: $\mathbb{Z}_3$ structure would fail
\end{itemize}
\end{tcolorbox}

% ------------------------------------------------------------------------------
% DEPENDENCY & STATUS (IF/THEN structure)
% ------------------------------------------------------------------------------

\begin{tcolorbox}[colback=blue!5!white, colframe=blue!50!black,
    title=\textbf{Dependency \& Status}]
\textbf{Inputs required from earlier chapters:}
\begin{itemize}[nosep]
    \item Ch.~\ref{ch:three_generations}: $\mathbb{Z}_6 = \mathbb{Z}_2 \times \mathbb{Z}_3$ lattice symmetry
    \item Ch.~\ref{ch:va_structure}: Boundary conditions $\Rightarrow$ chirality selection
    \item Ch.~\ref{ch:gf_derivation}: Fermi constant mechanism (overlap integral formalism)
\end{itemize}

\textbf{IF/THEN structure:}
\begin{itemize}[nosep]
    \item \textbf{IF} $\mathbb{Z}_3$ structure established \textbf{THEN} three neutrino flavors follow \tagI{}
    \item \textbf{IF} Higgs localized in interior, $\nu$ at interface \textbf{THEN} $m_\nu/m_e \sim e^{-14}$ \tagDc{}
    \item \textbf{IF} boundary conditions select $L$ \textbf{THEN} only $\nu_L$ couples \tagDc{}
    \item \textbf{IF} $\mathbb{Z}_6$ discrete overlap computed \textbf{THEN} $\theta_{23} \approx 45°$ \tagDc{}
\end{itemize}

\textbf{Status:} YELLOW --- Mechanism sound, profiles postulated, angles partially derived.
\end{tcolorbox}

% ------------------------------------------------------------------------------
% BOOK-READY INTRODUCTION
% ------------------------------------------------------------------------------

\paragraph{Chapter overview.}
Neutrinos are the lightest known fermions, with masses at least six orders of
magnitude below the electron. In the Standard Model, this hierarchy is an
unexplained input. EDC offers a geometric explanation: neutrinos are
\emph{edge modes}---boundary excitations localized at the interface between
bulk and brane, with suppressed overlap to the Higgs/mass mechanism residing
in the brane interior. This overlap suppression naturally produces $m_\nu \ll m_e$
without fine-tuning \tagDc{}.

For flavor mixing (PMNS matrix), EDC provides a \emph{computed baseline}: if the
three neutrino flavors correspond to $\mathbb{Z}_3$ modes, the simplest symmetric
assumption yields the discrete Fourier transform (DFT) matrix with all
$|U_{\alpha i}|^2 = 1/3$. This baseline predicts $\sin^2\theta_{13} = 1/3$,
which is \textbf{15 times larger} than the observed value of $\approx 0.022$.
The DFT baseline is therefore \emph{falsified}, indicating that $\mathbb{Z}_3$
symmetry must be broken at the $\sim 25\%$ level in the $\nu_e$--$\nu_3$ sector.
This is a tight negative result that closes the logical loop and identifies the
required physics \tagDc{}.

% ------------------------------------------------------------------------------
% READER MAP
% ------------------------------------------------------------------------------

\begin{tcolorbox}[colback=blue!5, colframe=blue!50!black,
    title=\textbf{Reader Map: What This Chapter Establishes}]
\begin{description}[style=nextline, leftmargin=1em, font=\normalfont\bfseries]
    \item[Derived \tagDc{}:]
        Mass suppression via overlap integrals (if profiles given);
        left-handed selection from Chapter~\ref{ch:va_structure} boundary conditions;
        $\mathbb{Z}_3$ DFT baseline for PMNS.

    \item[Identified \tagI{}:]
        Three neutrino flavors $\leftrightarrow$ $|\mathbb{Z}_3| = 3$;
        mass hierarchy $\leftrightarrow$ mode number.

    \item[Postulated \tagP{}:]
        Edge-mode ontology;
        overlap-based PMNS mechanism;
        specific breaking of $\mathbb{Z}_3$ (mechanism not derived).

    \item[Open (not addressed):]
        Absolute mass scale;
        explicit $\kappa^{-1}$ from EDC action;
        PMNS angles from breaking;
        CP phase $\delta$;
        Dirac vs.\ Majorana nature.
\end{description}
\end{tcolorbox}

% ------------------------------------------------------------------------------
% MINI SUMMARY TABLE
% ------------------------------------------------------------------------------

\begin{table}[ht]
\centering
\caption{Chapter 6 mechanism summary}
\label{tab:ch6_mechanism_summary}
\begin{tabular}{p{3.5cm}p{3.5cm}p{3.5cm}c}
\toprule
\textbf{Mechanism} & \textbf{Inputs} & \textbf{Output} & \textbf{Tag} \\
\midrule
Edge-mode localization & Neutrino at interface \tagP{} &
    Suppressed Higgs overlap & \tagDc{} \\
Overlap $\to$ mass & Profiles $f_\nu(\xi)$, $h(\xi)$ \tagP{} &
    $m_\nu/m_e \sim e^{-\Delta \xi/\kappa^{-1}}$ & \tagDc{} \\
$\mathbb{Z}_3$ flavor count & Hexagonal $\mathbb{Z}_6 = \mathbb{Z}_2 \times \mathbb{Z}_3$ &
    $N_\nu = 3$ & \tagI{} \\
DFT baseline (PMNS) & $\mathbb{Z}_3$ symmetric mass &
    $|U_{\alpha i}|^2 = 1/3$ & \tagDc{} \\
DFT vs.\ PDG & $\sin^2\theta_{13}^{\text{DFT}} = 0.333$ &
    \textbf{Falsified} ($\times 15$ off) & --- \\
Breaking requirement & Falsified baseline &
    $\sim 25\%$ anisotropy needed & \tagP{} \\
\bottomrule
\end{tabular}
\end{table}

% ==============================================================================
\subsection{The Neutrino Problem}
\label{sec:ch6_problem}

Neutrino physics presents several puzzles that demand explanation:

\begin{table}[ht]
\centering
\caption{Neutrino baseline facts \tagBL{}}
\label{tab:ch6_baselines}
\begin{tabular}{lll}
\toprule
\textbf{Observable} & \textbf{Value (PDG 2024)} & \textbf{Puzzle} \\
\midrule
Absolute mass & $m_\nu \lesssim 0.8$ eV (direct) & Why $m_\nu/m_e \sim 10^{-6}$? \\
$\Delta m_{21}^2$ & $7.53 \times 10^{-5}$ eV$^2$ & Why this splitting? \\
$|\Delta m_{31}^2|$ & $2.453 \times 10^{-3}$ eV$^2$ & Why hierarchical? \\
Weak coupling & Only left-handed couple & Why chirality-selected? \\
Three flavors & $N_\nu = 2.984 \pm 0.008$ (LEP) & Why exactly three? \\
\bottomrule
\end{tabular}
\end{table}

In the Standard Model, these are input parameters with no deeper explanation.
EDC proposes a geometric origin.

% ==============================================================================
\subsection{Toy Model: The Boundary-Trapped State}
\label{sec:ch6_toy_model}

Before diving into the formal framework, consider a minimal intuition model
that captures the essential physics.

\paragraph{The setup.}
Imagine a 1D potential $V(\xi)$ with the following structure:
\begin{itemize}[nosep]
    \item For $\xi < 0$ (bulk): $V(\xi) = V_{\text{bulk}} > 0$ (large positive value)
    \item At $\xi = 0$ (interface): sharp boundary where potential changes
    \item For $\xi > 0$ (brane interior): $V(\xi) = 0$ (attractive region)
\end{itemize}

A particle in this potential has two types of states:
\begin{enumerate}[nosep]
    \item \textbf{Interior modes:} Wavefunction peaked at $\xi > 0$, freely propagating
          in the brane. These are electrons, quarks---``normal'' particles.
    \item \textbf{Edge mode:} Wavefunction trapped at $\xi \approx 0$, exponentially
          decaying into both bulk (blocked by $V_{\text{bulk}}$) and interior.
          This is the neutrino.
\end{enumerate}

\paragraph{Why the edge mode has small mass.}
Add a ``Higgs bump'' $h(\xi)$ that is nonzero only for $\xi > \xi_H$ (deep in the interior).
An interior mode overlaps fully with $h(\xi)$ and gets mass $m \sim m_0$.
The edge mode, trapped at $\xi = 0$, overlaps only via its exponential tail:
$m_{\text{edge}} \sim m_0 \cdot e^{-\xi_H / \lambda}$ where $\lambda$ is the decay length.

\paragraph{What this toy model captures:}
\begin{itemize}[nosep]
    \item[\ding{51}] Why $m_\nu \ll m_e$: edge mode has suppressed Higgs overlap
    \item[\ding{51}] Why only three: if the boundary has $\mathbb{Z}_3$ angular structure,
          there are three inequivalent edge positions
    \item[\ding{51}] Why left-handed only: boundary conditions at $\xi = 0$ select chirality
          (see Ch.~\ref{ch:va_structure})
\end{itemize}

\paragraph{What this toy model ignores:}
\begin{itemize}[nosep]
    \item[$\times$] The actual form of $V(\xi)$ from EDC action (not derived: OPR-04)
    \item[$\times$] Angular structure of the edge mode (PMNS mixing requires this)
    \item[$\times$] Absolute mass scale (requires $\lambda$, $z_H$ values)
    \item[$\times$] Mass hierarchy between $\nu_1, \nu_2, \nu_3$
\end{itemize}

\begin{tcolorbox}[colback=yellow!5!white, colframe=yellow!60!black,
    title=\textbf{Toy Model Status}]
This 1D boundary-trapped picture is \textbf{pedagogical} \tagP{}, not derived.
It correctly predicts:
\begin{itemize}[nosep]
    \item Exponential mass suppression (structure)
    \item Chirality selection (from boundary)
    \item Three flavors (from $\mathbb{Z}_3$ angular structure)
\end{itemize}
It does \textbf{not} predict:
\begin{itemize}[nosep]
    \item Numerical values of masses or mixing angles
    \item Which hierarchy (normal vs inverted)
    \item Dirac vs Majorana nature
\end{itemize}
The toy model is useful for \emph{understanding} the mechanism, not for
\emph{computing} predictions.
\end{tcolorbox}

% ------------------------------------------------------------------------------
% FIGURE PLACEHOLDERS
% ------------------------------------------------------------------------------

\begin{tcolorbox}[colback=gray!10, colframe=gray!50, title=\textbf{Figure Placeholder 1: Edge-Mode Localization Schematic}]
\textbf{Suggested content:}
\begin{itemize}[nosep]
    \item Side view of thick brane showing $\xi$-axis (fifth dimension)
    \item Three regions labeled: 5D Bulk (left), Interface (center), Brane Interior (right)
    \item Neutrino wavefunction $|\psi_\nu(\xi)|^2$ peaked at interface, decaying both ways
    \item Electron wavefunction $|\psi_e(\xi)|^2$ peaked in interior
    \item Higgs profile $h(\xi)$ localized in interior (shaded region)
    \item Overlap region highlighted: neutrino tail meets Higgs $\Rightarrow$ tiny $m_\nu$
\end{itemize}
\textbf{Key message:} Spatial separation between edge mode ($\nu$) and Higgs ($h$) causes mass suppression.
\end{tcolorbox}

\begin{tcolorbox}[colback=gray!10, colframe=gray!50, title=\textbf{Figure Placeholder 2: PMNS Mixing / Angular Overlap Intuition}]
\textbf{Suggested content:}
\begin{itemize}[nosep]
    \item Top view of hexagonal lattice showing $\mathbb{Z}_6$ angular structure
    \item Three flavor sectors marked at $\phi = 0, 2\pi/3, 4\pi/3$ (neutrino flavors)
    \item Three mass eigenstates as radial modes with different penetration depths
    \item Overlap regions between flavor sectors and mass eigenstates
    \item Large overlap ($\theta_{23} \approx 45°$) vs small overlap ($\theta_{13} \approx 8.5°$) visualized
\end{itemize}
\textbf{Key message:} PMNS angles reflect angular overlap between flavor and mass eigenstates in $\mathbb{Z}_6$ geometry.
\end{tcolorbox}

% ==============================================================================
\subsection{Edge-Mode Ontology}
\label{sec:ch6_ontology}

\subsubsection{The Neutrino as Boundary Excitation}

\begin{postulate}[Neutrino as Edge Mode {\normalfont \tagP{}}]
\label{post:ch6_neutrino_edge}
The neutrino is an \textbf{edge mode}---a boundary excitation localized at the
interface between the 5D bulk and the 3D brane. Unlike interior brane modes
(electron, quarks) or bulk-penetrating modes (proton junction), the neutrino
wavefunction peaks at the interface:
\begin{equation}
    |\psi_\nu(\xi)|^2 \propto e^{-2\kappa |\xi - \xi_{\text{interface}}|}
    \label{eq:ch6_edge_profile}
\end{equation}
where $\kappa^{-1}$ is the penetration depth into the brane interior.
\end{postulate}

\paragraph{Physical picture.}
The thick brane has three conceptual layers:
\begin{enumerate}[nosep]
    \item \textbf{Bulk} ($\xi < 0$): 5D Plenum region
    \item \textbf{Interface} ($\xi \approx 0$): Transition zone where boundary conditions apply
    \item \textbf{Brane interior} ($\xi > 0$): Where charged leptons and quarks localize
\end{enumerate}
The neutrino resides in layer 2---the interface---with exponentially suppressed
coupling to both bulk and interior.

\begin{center}
\begin{tikzpicture}[scale=0.85]
  % Bulk region
  \fill[blue!10] (-4,-2.5) rectangle (4,-1.5);
  \node[font=\scriptsize] at (0,-2) {5D Bulk (Plenum)};

  % Interfacial zone (neutrino region)
  \fill[purple!15] (-4,-1.5) rectangle (4,-0.7);
  \node[font=\scriptsize, purple!70!black] at (-2.5,-1.1) {Interface (edge modes)};

  % Brane internal layer
  \fill[yellow!20] (-4,-0.7) rectangle (4,0.7);
  \node[font=\scriptsize] at (-2.5,0) {Brane interior};

  % Observer region
  \fill[green!10] (-4,0.7) rectangle (4,1.7);
  \node[font=\scriptsize] at (0,1.2) {3D Observer space};

  % Neutrino (interface)
  \fill[purple!70] (0,-1.1) circle (0.18);
  \node[font=\scriptsize\bfseries, right] at (0.25,-1.1) {$\nu$};

  % Electron (interior)
  \fill[red!70] (2,0.4) circle (0.15);
  \node[font=\tiny, right] at (2.2,0.4) {$e^-$};

  % Boundaries
  \draw[thick, dashed] (-4,-1.5) -- (4,-1.5);
  \draw[thick, dashed, purple] (-4,-0.7) -- (4,-0.7);
  \draw[thick] (-4,0.7) -- (4,0.7);

  % z-axis
  \draw[->, thick] (4.5,-2.5) -- (4.5,1.7);
  \node[font=\scriptsize, right] at (4.5,1.7) {$\xi$};
\end{tikzpicture}
\end{center}

\subsubsection{Distinction from ``Bulk Escape''}

\begin{tcolorbox}[edcWarning, title={Language Precision}]
\textbf{Avoid:} ``The neutrino escapes into the bulk and cannot be detected.''

\textbf{Use:} ``The neutrino is an edge mode with \textbf{suppressed overlap}
to observer-facing states.'' \tagP{}

The first statement implies energy loss to extra dimensions, violating observed
4D energy conservation. The second correctly attributes weak coupling to
geometric suppression.
\end{tcolorbox}

% ==============================================================================
\subsection{Mass Suppression Mechanism}
\label{sec:ch6_mass_suppression}

\subsubsection{The Overlap Integral Argument}

The effective 4D mass of a fermion arises from overlap integrals over the
fifth dimension \tagBL{}:
\begin{equation}
    m_{\text{eff}} \sim m_0 \int_{-\infty}^{\infty} |f(\xi)|^2 \, h(\xi) \, d\xi
    \label{eq:ch6_overlap}
\end{equation}
where:
\begin{itemize}[nosep]
    \item $f(\xi)$ is the fermion's $\xi$-profile
    \item $h(\xi)$ is the Higgs/mass-generating field profile (brane-localized)
    \item $m_0$ is the 5D mass scale
\end{itemize}

\subsubsection{Why Neutrino Mass is Suppressed}

If the Higgs profile $h(\xi)$ is localized in the brane interior ($\xi > 0$),
but the neutrino profile $\psi_\nu(\xi)$ peaks at the interface ($\xi \approx 0$):

\begin{proposition}[Mass Suppression {\normalfont \tagDc{}}]
\label{prop:ch6_suppression}
The ratio of neutrino to electron mass is exponentially suppressed by
spatial separation:
\begin{equation}
    \frac{m_\nu}{m_e} \sim \exp\left(-\frac{\Delta z}{\kappa^{-1}}\right)
    \label{eq:ch6_mass_ratio}
\end{equation}
where $\Delta z$ is the separation between the neutrino interface position
and the Higgs localization, and $\kappa^{-1}$ is the neutrino penetration depth.
\end{proposition}

\begin{proof}[Derivation \tagDc{}]
For the electron (interior mode) with profile $f_e(z)$ peaked at $\xi = z_H$:
\begin{equation}
    m_e \sim m_0 \int |f_e(\xi)|^2 h(\xi) \, d\xi \approx m_0 \cdot 1
\end{equation}
(normalized overlap).

For the neutrino (edge mode) with profile peaked at $\xi = 0$:
\begin{equation}
    m_\nu \sim m_0 \int |f_\nu(\xi)|^2 h(\xi) \, d\xi
    \approx m_0 \cdot e^{-2\kappa z_H}
\end{equation}
since $|f_\nu(z_H)|^2 \approx e^{-2\kappa z_H}$.

The ratio follows:
\begin{equation}
    \frac{m_\nu}{m_e} \approx e^{-2\kappa z_H} = e^{-\Delta z/\kappa^{-1}}
    \quad \text{with } \Delta z \equiv 2\kappa z_H
\end{equation}
\end{proof}

\paragraph{Numerical estimate.}
For $m_\nu/m_e \sim 10^{-6}$, we need:
\begin{equation}
    e^{-\Delta z / \kappa^{-1}} \sim 10^{-6}
    \quad\Longrightarrow\quad
    \frac{\Delta z}{\kappa^{-1}} \approx 14
\end{equation}
This is geometrically reasonable: the separation is $\sim 14$ penetration depths.

\subsubsection{Stoplight Verdict: Mass Suppression}

\begin{table}[ht]
\centering
\caption{Mass suppression mechanism audit}
\label{tab:ch6_mass_stoplight}
\begin{tabular}{lccc}
\toprule
\textbf{Step} & \textbf{Status} & \textbf{Tag} & \textbf{Issue} \\
\midrule
Overlap integral formalism & GREEN & \tagBL{} & Standard KK reduction \\
Higgs localized in interior & YELLOW & \tagP{} & Profile not derived \\
Neutrino at interface & YELLOW & \tagP{} & Ontology postulated \\
Exponential suppression & GREEN & \tagDc{} & Follows from profiles \\
$m_\nu/m_e \sim 10^{-6}$ & YELLOW & \tagI{} & Requires $\Delta z/\kappa^{-1} \approx 14$ \\
Absolute $m_\nu$ value & RED & (open) & Not computed (OPR-04) \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Verdict: YELLOW} --- The suppression mechanism is geometrically sound
\tagDc{}, but profile shapes are postulated \tagP{}, and absolute masses are
not derived.

% ==============================================================================
\subsection{Three Neutrino Flavors}
\label{sec:ch6_three_flavors}

\subsubsection{Connection to Generation Structure}

Chapter~\ref{ch:three_generations} identified the generation count $N_{\text{gen}} = 3$
with the $\mathbb{Z}_3$ factor in the hexagonal lattice symmetry:
\begin{equation}
    \mathbb{Z}_6 = \mathbb{Z}_2 \times \mathbb{Z}_3
    \quad\Longrightarrow\quad
    |\mathbb{Z}_3| = 3
    \label{eq:ch6_z3}
\end{equation}

\begin{postulate}[Three Neutrino Flavors {\normalfont \tagI{}}]
\label{post:ch6_three_nu}
The three neutrino flavors $(\nu_e, \nu_\mu, \nu_\tau)$ correspond to the
three elements of $\mathbb{Z}_3$:
\begin{equation}
    \nu_i \leftrightarrow \omega^i, \qquad \omega = e^{2\pi i/3}, \quad i = 0, 1, 2
\end{equation}
Each flavor is an edge mode with angular quantum number $n = i$ around the
hexagonal axis.
\end{postulate}

This is an \textbf{identification} \tagI{}, not a derivation---the $\mathbb{Z}_3$
structure provides the cardinality, but the dynamical mechanism connecting
neutrino wavefunctions to $\mathbb{Z}_3$ rotations is not derived.

\subsubsection{Mass Hierarchy from Mode Number}

By analogy with the charged lepton mass hierarchy (Chapter~\ref{ch:lepton_candidates}),
the neutrino masses may scale with mode number:
\begin{equation}
    m_{\nu_i} \propto f(n_i) \cdot e^{-\Delta z_i/\kappa^{-1}}
    \label{eq:ch6_nu_hierarchy}
\end{equation}
where $n_i \in \{0, 1, 2\}$ labels the generation.

\paragraph{Status:} This scaling is \textbf{postulated} \tagP{}. The function $f(n)$
and the dependence of $\Delta z$ on mode number are not derived.

% ==============================================================================
\subsection{Connection to V--A Chirality}
\label{sec:ch6_va_connection}

Chapter~\ref{sec:ch9_va_structure} derived that only left-handed fermions couple
at the brane interface \tagDc{}. This applies directly to neutrinos:

\begin{corollary}[Left-Handed Neutrinos {\normalfont \tagDc{}}]
\label{cor:ch6_left_nu}
The boundary conditions that select left-handed charged leptons (Ch.~9)
simultaneously select left-handed neutrinos:
\begin{equation}
    P_L \psi_\nu = \psi_{\nu,L} \quad \text{(normalizable edge mode)}
\end{equation}
Right-handed neutrinos, if they exist, are expelled into the bulk and do not
couple to the weak vertex.
\end{corollary}

\paragraph{Consistency check.}
The observed V--A structure of weak currents \tagBL{}:
\begin{equation}
    \mathcal{J}^\mu_{\text{weak}} = \bar\psi_\ell \gamma^\mu (1 - \gamma^5) \psi_\nu
\end{equation}
emerges from the same boundary-condition mechanism that produces chiral
localization (Ch.~9), applied to the neutrino edge mode. No additional
assumptions are required.

% ==============================================================================
\subsection{PMNS Mixing: Postulated Structure}
\label{sec:ch6_pmns}

\subsubsection{The Mixing Matrix}

Neutrino flavor eigenstates $(\nu_e, \nu_\mu, \nu_\tau)$ are related to
mass eigenstates $(\nu_1, \nu_2, \nu_3)$ by the PMNS matrix \tagBL{}:
\begin{equation}
    \begin{pmatrix} \nu_e \\ \nu_\mu \\ \nu_\tau \end{pmatrix}
    = U_{\text{PMNS}}
    \begin{pmatrix} \nu_1 \\ \nu_2 \\ \nu_3 \end{pmatrix}
    \label{eq:ch6_pmns}
\end{equation}

The observed mixing angles (PDG 2024) \tagBL{}:
\begin{align}
    \sin^2\theta_{12} &\approx 0.307 \quad (\text{solar}) \\
    \sin^2\theta_{23} &\approx 0.546 \quad (\text{atmospheric}) \\
    \sin^2\theta_{13} &\approx 0.022 \quad (\text{reactor})
\end{align}

\subsubsection{EDC Interpretation (Postulated)}

\begin{postulate}[PMNS from Wavefunction Overlap {\normalfont \tagP{}}]
\label{post:ch6_pmns}
The PMNS mixing arises from overlap integrals between flavor wavefunctions
(edge modes at different $\mathbb{Z}_3$ angles) and mass wavefunctions
(Higgs-coupled modes):
\begin{equation}
    (U_{\text{PMNS}})_{\alpha i} \propto \int \psi^*_{\nu_\alpha}(\xi, \phi) \,
    \psi_{\nu_i}^{\text{mass}}(\xi, \phi) \, d\xi \, d\phi
    \label{eq:ch6_pmns_overlap}
\end{equation}
where $\phi$ is the angular coordinate around the $\mathbb{Z}_6$ axis.
\end{postulate}

\textbf{Status: RED} \tagP{} --- This is a structural postulate. No explicit
calculation of PMNS angles from EDC geometry has been performed.

\subsubsection{Stoplight Verdict: PMNS Mixing}

\begin{table}[ht]
\centering
\caption{PMNS mixing mechanism audit}
\label{tab:ch6_pmns_stoplight}
\begin{tabular}{lccc}
\toprule
\textbf{Claim} & \textbf{Status} & \textbf{Tag} & \textbf{Issue} \\
\midrule
$U_{\text{PMNS}}$ exists & GREEN & \tagBL{} & Observed \\
Mixing from overlaps & RED & \tagP{} & Mechanism not computed (OPR-05) \\
$\theta_{12} \approx 33°$ & RED & (open) & Not derived (OPR-05) \\
$\theta_{23} \approx 45°$ & RED & (open) & Not derived (OPR-05) \\
$\theta_{13} \approx 8.5°$ & RED & (open) & Not derived (OPR-05) \\
CP phase $\delta$ & RED & (open) & Not addressed (OPR-06) \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Verdict: RED} --- PMNS structure is postulated, not derived.

% ------------------------------------------------------------------------------
% Include the PMNS symmetry baseline calculation (Attempt 1)
% ------------------------------------------------------------------------------
\input{sections/ch6_pmns_attempt1}

% ------------------------------------------------------------------------------
% Include the PMNS overlap model (Attempt 2)
% ------------------------------------------------------------------------------
\input{sections/ch6_pmns_attempt2}

% ------------------------------------------------------------------------------
% Include the Z6 discrete-phase refinement (Attempt 3)
% ------------------------------------------------------------------------------
\input{sections/ch6_pmns_attempt3_z6_refinement}

% ------------------------------------------------------------------------------
% Include the rank-2 and double-path mechanisms (Attempt 4)
% ------------------------------------------------------------------------------
\input{sections/ch6_pmns_attempt4_menu}

% ------------------------------------------------------------------------------
% Include the epsilon derivation (Attempt 4.1)
% ------------------------------------------------------------------------------
\input{sections/ch6_pmns_attempt4_1_derive_epsilon}

% ------------------------------------------------------------------------------
% Include the theta12 origin micro-attempt (Attempt 4.2)
% ------------------------------------------------------------------------------
\input{sections/ch6_pmns_attempt4_2_theta12_origin}

% ==============================================================================
\subsection{Dirac vs. Majorana Nature}
\label{sec:ch6_dirac_majorana}

\subsubsection{The Open Question}

Whether neutrinos are Dirac or Majorana particles remains experimentally
undetermined \tagBL{}. The key observable is neutrinoless double-beta decay
($0\nu\beta\beta$):
\begin{itemize}[nosep]
    \item If observed: neutrinos are Majorana
    \item If not observed (with sufficient sensitivity): inconclusive
\end{itemize}

\subsubsection{EDC Perspective}

\begin{remark}[Edge-Mode Nature {\normalfont \tagP{}}]
In the edge-mode ontology, the Dirac/Majorana distinction maps to:
\begin{itemize}[nosep]
    \item \textbf{Dirac:} Distinct $\nu$ and $\bar\nu$ edge modes with opposite
          quantum numbers
    \item \textbf{Majorana:} Single self-conjugate edge mode where ``antineutrino''
          is the same mode with opposite helicity
\end{itemize}
EDC does not currently distinguish these cases; both are compatible with the
edge-mode picture.
\end{remark}

\textbf{Status:} (open) --- The edge-mode framework accommodates both, but
makes no prediction.

% ------------------------------------------------------------------------------
% FAILURE MODES
% ------------------------------------------------------------------------------

\begin{tcolorbox}[colback=red!5!white, colframe=red!50!black,
    title=\textbf{Failure Modes: What Could Go Wrong}]
\begin{description}[style=nextline, leftmargin=1em, font=\normalfont\bfseries]
    \item[FM-1: No edge-mode solution exists.]
        The thick-brane BVP (Ch.~\ref{sec:ch12_bvp_workpackage}) might have no
        boundary-localized eigenstate for physical parameters.
        \emph{Implication:} Edge-mode ontology fails; neutrino must be explained differently.

    \item[FM-2: Overlap suppression insufficient.]
        If $\Delta z / \kappa^{-1} \ll 14$, the mass ratio $m_\nu/m_e$ would be
        too large (e.g., $10^{-3}$ instead of $10^{-6}$).
        \emph{Implication:} Profile parameters need adjustment, or mechanism is wrong.

    \item[FM-3: Wrong chirality selection.]
        Boundary conditions might select $\nu_R$ instead of $\nu_L$, contradicting
        observed V$-$A structure.
        \emph{Implication:} BC ansatz in Ch.~\ref{ch:va_structure} incorrect.

    \item[FM-4: PMNS angles incompatible.]
        The $\mathbb{Z}_6$ overlap structure might predict angles that cannot be
        corrected to match observation (e.g., $\theta_{12} = 0$ forced).
        \emph{Implication:} Additional breaking mechanism required.

    \item[FM-5: Fourth neutrino discovered.]
        LEP already constrains $N_\nu = 2.984 \pm 0.008$, but future precision
        could reveal a light sterile neutrino.
        \emph{Implication:} $\mathbb{Z}_3$ identification fails; need $\mathbb{Z}_4$ or continuous.
\end{description}

\textbf{Current status:} No failure mode triggered. All observations consistent with framework.
\end{tcolorbox}

% ==============================================================================
\subsection{Summary and Falsifiability}
\label{sec:ch6_summary}

\subsubsection{What Chapter 6 Establishes}

\begin{tcolorbox}[colback=green!5, colframe=green!50!black,
    title=\textbf{Chapter 6 Summary}]
\begin{enumerate}[nosep]
    \item \textbf{Mass suppression:} Neutrino smallness ($m_\nu \ll m_e$)
          explained by edge-mode localization and suppressed overlap with
          Higgs profile \tagDc{}
    \item \textbf{Chirality:} Left-handed selection follows from Ch.~9
          boundary conditions \tagDc{}
    \item \textbf{Three flavors:} Connected to $\mathbb{Z}_3$ structure
          (Ch.~5), but dynamical mechanism not derived \tagI{}
    \item \textbf{PMNS $\theta_{23}$:} Derived from $\mathbb{Z}_6$ geometry
          (Attempt~2); preserved in rank-2 construction (Attempt~4) \tagDc{}
    \item \textbf{PMNS $\theta_{12}$, $\theta_{13}$:} Structure identified
          (Attempt~4: rank-2 + reactor perturbation); values require geometric derivation \tagI{}
\end{enumerate}
\end{tcolorbox}

\subsubsection{What Remains Open}

\begin{table}[ht]
\centering
\caption{Open problems in EDC neutrino physics}
\label{tab:ch6_open}
\begin{tabular}{p{5.5cm}p{5.5cm}}
\toprule
\textbf{Open Problem} & \textbf{Required Progress} \\
\midrule
Absolute neutrino mass scale & Derive $\kappa^{-1}$ and $\Delta z$ from EDC action \\
Mass hierarchy (normal/inverted) & Connect to mode spectrum \\
PMNS angles from geometry & Explicit overlap calculation \\
CP violation phase $\delta$ & Requires complex phase mechanism \\
Dirac vs. Majorana & Edge-mode self-conjugacy condition \\
Sterile neutrino coupling & Interface boundary condition modification \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Falsifiability Clause}

\begin{tcolorbox}[edcWarning, title={Falsification Criteria}]
The EDC neutrino-as-edge-mode hypothesis would be \textbf{falsified} if:
\begin{enumerate}[nosep]
    \item \textbf{Right-handed weak coupling:} Right-handed neutrinos observed
          coupling to $W^\pm$ with comparable strength to left-handed
    \item \textbf{Large neutrino mass:} $m_\nu > 1$ eV directly measured
          (would require much smaller separation, inconsistent with
          $m_e$ explanation)
    \item \textbf{Stronger interactions:} Neutrino cross-sections significantly
          larger than geometric overlap predicts
    \item \textbf{Bulk propagation:} Modified dispersion relation at high
          energy indicating bulk escape
    \item \textbf{4th light neutrino:} Discovery of a fourth sequential
          neutrino coupling to $Z$ (would falsify $\mathbb{Z}_3$ count)
\end{enumerate}
Current data: All observations consistent with EDC predictions \tagBL{}.
\end{tcolorbox}

\subsubsection{Overall Stoplight}

\begin{table}[ht]
\centering
\caption{Chapter 6 overall verdict}
\label{tab:ch6_verdict}
\begin{tabular}{lcc}
\toprule
\textbf{Claim} & \textbf{Verdict} & \textbf{Tag} \\
\midrule
$m_\nu \ll m_e$ from geometry & YELLOW & \tagDc{} \\
$N_\nu = 3$ from $\mathbb{Z}_3$ & YELLOW & \tagI{} \\
Left-handed selection & GREEN & \tagDc{} \\
PMNS: $\mathbb{Z}_3$ DFT baseline (Attempt 1) & FALSIFIED & \tagDc{} \\
PMNS: $\mathbb{Z}_6$ overlap (Attempt 2) & YELLOW & \tagDc{} (OPR-05) \\
PMNS: $\mathbb{Z}_6$ discrete phases (Attempt 3) & RED & \tagDc{} \\
PMNS: Rank-2 + $\varepsilon$ (Attempt 4) & \textcolor{green!60!black}{\textbf{GREEN*}} & \tagI{} (OPR-05) \\
$\theta_{23}$ from geometry & GREEN & \tagDc{} \\
$\theta_{12}, \theta_{13}$ structure & YELLOW & \tagI{} (values not derived) \\
Dirac/Majorana prediction & RED & (open) (OPR-07) \\
\bottomrule
\end{tabular}

\vspace{0.5em}
{\footnotesize *GREEN numerically; YELLOW epistemically (see Attempt~4 assessment).}
\end{table}

\textbf{Bottom line:} EDC provides a structural explanation for neutrino
smallness and chirality \tagDc{}, an identification of the flavor count with
$\mathbb{Z}_3$ \tagI{}, and four computed PMNS approaches: the DFT baseline (Attempt~1)
is falsified; the $\mathbb{Z}_6$ overlap model (Attempt~2) \emph{derives} the
atmospheric mixing angle $\theta_{23} \approx 45°$ from pure geometry \tagDc{};
Attempt~3 shows that discrete Z$_6$ phases alone cannot correct the remaining angles;
and Attempt~4 identifies the \emph{structure} (rank-2 baseline + reactor perturbation)
that reproduces all three PMNS angles within 3\%, though the values of $\theta_{12}^0$
and $\varepsilon$ require geometric derivation \tagI{}. Dirac/Majorana nature remains open.

