% ============================================================================
\section{Introduction}
\label{sec:intro}
% ============================================================================

\subsection{Purpose of This Document}

This companion paper provides formal definitions for the symmetry structures and topological operations used in Elastic Diffusive Cosmology (EDC). It serves as a reference document that the main paper~\cite{paper3} (and subsequent EDC publications) can cite without repeating foundational material.

The content here does \emph{not} claim to derive new physics. Rather, it organizes and formalizes structures that arise naturally from the EDC framework, assigns consistent epistemic tags, and identifies falsifiability conditions.

\subsection{The Two-Sided Reading Rule}

Throughout this document, we maintain a strict distinction between two levels of description:

\begin{itemize}[leftmargin=*]
    \item \textbf{5D cause} (geometric mechanism): Statements about the topology, geometry, and dynamics of defects in the 5D manifold $\Mfive$.
    \item \textbf{4D evidence} (observable prediction): Statements about quantities measurable in our 4D spacetime---masses, charges, decay rates, cross-sections.
\end{itemize}

\noindent The 5D cause is the \emph{proposal}; the 4D evidence is the \emph{test}. Neither side alone constitutes a complete claim. For example:
\begin{quote}
\emph{5D cause}: ``The $\beta^-$ decay corresponds to a $\Zsix$ sector shift in the Y-junction configuration.''\\
\emph{4D evidence}: ``The neutron lifetime and decay products match SM predictions.''
\end{quote}

\subsection{Relationship to Other EDC Documents}

This paper assumes familiarity with:
\begin{enumerate}
    \item \textbf{EDC Framework Reference} (v2.0): Defines the 5D manifold, action, and defect taxonomy.
    \item \textbf{Paper 3}: Uses the operators and ledger defined here for neutron physics.
\end{enumerate}

\noindent We do not duplicate content from the Framework Reference; instead, we cite it and build upon its definitions.

\subsection{Document Structure}

\begin{enumerate}
    \item \textbf{Section~\ref{sec:geometric-setup}}: Minimal geometric setup (manifold, parameters)
    \item \textbf{Section~\ref{sec:symmetry-layering}}: Symmetry layering (kinematic + internal)
    \item \textbf{Section~\ref{sec:defect-classification}}: Defect classification as layered structure
    \item \textbf{Section~\ref{sec:process-operators}}: Topological process operators ($\opE$, $\opR$, $\opM$)
    \item \textbf{Section~\ref{sec:beta-ledger}}: $\beta^-$ decay conservation ledger
    \item \textbf{Section~\ref{sec:discussion}}: Discussion and research roadmap
\end{enumerate}
