% ==============================================================================
% Chapter 7, Attempt 3: CP Phase from Z3 Structure
% Status: FALSIFIED — Phase Cancellation Theorem (Attempt 4) shows pure Z3 gives J=0
% Retained for historical record; Attempt 4 is the active backbone
% ==============================================================================

\subsection{Attempt 3: CP Phase and the Jarlskog Invariant}
\label{sec:ch7_attempt3}

% ------------------------------------------------------------------------------
% FALSIFIED BANNER — This attempt is superseded by Attempt 4
% ------------------------------------------------------------------------------
\begin{tcolorbox}[colback=red!15, colframe=red!70!black,
    title=\textbf{STATUS: FALSIFIED --- Inconsistent Phase Ansatz}]
\textbf{This attempt is superseded by Attempt~4.}

Attempt~3 explored whether $\mathbb{Z}_3$ discrete phases could produce CP violation.
The numerical result ($J \simeq 2.9 \times 10^{-5}$) appeared promising, but
\textbf{Attempt~4 proves this ansatz is inconsistent}:

\begin{itemize}[nosep]
    \item \textbf{Phase Cancellation Theorem (Attempt~4):} For \emph{any} $\mathbb{Z}_3$
          charge assignment to quark generations, the Jarlskog phase sum vanishes
          identically: $J = 0$.
    \item \textbf{Inconsistency:} The phase assignment used here
          ($\omega^0, \omega^{-2}, \omega^{-2}, \omega^{-1}$) does not correspond
          to any valid $\mathbb{Z}_3$ charge configuration.
    \item \textbf{Resolution:} Attempt~4 shows that the $\mathbb{Z}_2$ factor in
          $\mathbb{Z}_6 = \mathbb{Z}_2 \times \mathbb{Z}_3$ provides the necessary
          sign selection mechanism, yielding $\delta = 60°$ (5° from PDG).
\end{itemize}

\textbf{This section is retained as historical exploration only.}
The tables and analysis below document what was tested; \emph{none of it should be
cited as a valid result}. See Attempt~4 (Section~\ref{sec:ch7_attempt4}) for the
corrected analysis.
\end{tcolorbox}

\vspace{1em}

The overlap model (Attempts~2.1, 2.2) successfully derives the CKM magnitude
hierarchy ($\lambda$, $\lambda^2$, $\lambda^3$) from geometry. What remains
is the \textbf{complex phase structure}---specifically, the origin of CP
violation as measured by the Jarlskog invariant $J$ and the phase $\delta$.

At this stage the overlap model fixes the magnitude hierarchy; the remaining
task is to produce a \textbf{rephasing-invariant} complex phase (nonzero $J$),
i.e., physical CP violation.

\subsubsection{Quantitative Targets}

The Standard Model CP violation is characterized by \tagBL{}:
\begin{align}
    J &= \text{Im}(V_{us} V_{cb} V_{ub}^* V_{cs}^*) \approx 3.08 \times 10^{-5}
    \label{eq:ch7_J_pdg} \\
    \delta &\approx 1.14~\text{rad} \approx 65°
    \label{eq:ch7_delta_pdg} \\
    |\bar\rho - i\bar\eta| &\approx 0.38
    \label{eq:ch7_rhoeta_pdg}
\end{align}

For a proposed phase mechanism to produce \emph{physical} CP violation, it must
pass the \textbf{rephasing invariance test}: the phase must not be removable by
field redefinitions $V_{ij} \to e^{i\phi_i} V_{ij} e^{-i\phi_j'}$.

\subsubsection{Methodology: Two Tracks}

We systematically test six mechanisms under two constraints:

\begin{description}[style=nextline, leftmargin=1em]
    \item[\textbf{Track A} (no free parameters):]
        Use only quantities already fixed in Part~II: $\mathbb{Z}_6/\mathbb{Z}_3$
        structure, calibrated distances from Attempt~2.1, localization profiles.
        Success criterion: nonzero $J$ emerges without calibration.

    \item[\textbf{Track B} (one calibrated parameter):]
        Allow exactly one new parameter calibrated to one PDG input ($\delta$,
        $J$, or $|\bar\rho - i\bar\eta|$). Must produce at least one independent
        prediction not used in calibration.
\end{description}

\subsubsection{Option Menu}

Six mechanisms were tested (see \texttt{code/ckm\_cp\_attempt3.py} for details):

\begin{enumerate}
    \item[\textbf{O1:}] Complex z-shift / oscillatory bulk phase
    \item[\textbf{O2:}] Two-path interference (multiple overlap channels)
    \item[\textbf{O3:}] Boundary condition phase (MIT-bag type)
    \item[\textbf{O4:}] $\mathbb{Z}_6 = \mathbb{Z}_2 \times \mathbb{Z}_3$ discrete phases
    \item[\textbf{O5:}] Holonomy/Berry phase from $\mathbb{Z}_3$ cycle
    \item[\textbf{O6:}] Mediator-induced complex mixing
\end{enumerate}

\subsubsection{Track A Results}

\begin{table}[ht]
\centering
\caption{Track A: CP mechanisms with no free parameters}
\label{tab:ch7_trackA}
\begin{tabular}{llccl}
\toprule
\textbf{Option} & \textbf{Verdict} & \textbf{$J$ pred} & \textbf{$\delta$ pred} & \textbf{Reason} \\
\midrule
O1: Complex z-shift & RED & 0 & 0° & Phase removable by rephasing \\
O2: Two-path interf.\ & RED & 0 & 0° & No second path in framework \\
O3: Boundary phase & RED & 0 & 0° & Universal BC phase removable \\
\textbf{O4: $\mathbb{Z}_3$ phases} & \textbf{YELLOW} & $2.9 \times 10^{-5}$ & 120° & \textbf{J within 6\%!} \\
O5: Holonomy/Berry & YELLOW & $2.9 \times 10^{-5}$ & 120° & Equivalent to O4 \\
O6: Mediator mixing & RED & N/A & N/A & Mechanism undefined \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Key finding: Option 4 succeeds without calibration.}
The $\mathbb{Z}_3$ discrete symmetry, which already appears in Chapter~\ref{ch:three_generations}
for the generation count, naturally provides complex phases $\omega = e^{2\pi i/3}$
that survive in the rephasing-invariant combination $J$.

\subsubsection{The $\mathbb{Z}_3$ Phase Mechanism}

\begin{tcolorbox}[colback=blue!5, colframe=blue!50!black,
    title=\textbf{Option 4: How Z\textsubscript{3} Phases Generate CP Violation}]
\textbf{Setup:} Quark generations carry $\mathbb{Z}_3$ charges $q_i \in \{0, 1, 2\}$.
CKM elements acquire phases from the charge difference:
\begin{equation}
    V_{ij} \sim |V_{ij}| \cdot \omega^{q_i^{(u)} - q_j^{(d)}}, \qquad
    \omega = e^{2\pi i/3}
    \label{eq:ch7_z3_phase}
\end{equation}

\textbf{Rephasing check:} The Jarlskog invariant is:
\begin{equation}
    J = \text{Im}(V_{us} V_{cb} V_{ub}^* V_{cs}^*)
\end{equation}
Under rephasing $V_{ij} \to e^{i\phi_i} V_{ij} e^{-i\phi_j'}$, all phases cancel:
$J$ is \emph{invariant}. The $\mathbb{Z}_3$ phases contribute to $J$ non-trivially.

\textbf{Calculation:} With charge assignment $(0, 1, 2)$ for both up and down sectors,
the phase structure gives \tagDc{}:
\begin{align}
    V_{us} V_{cb} V_{ub}^* V_{cs}^* &\sim |...| \cdot \omega^{0-2+2+1} = |...| \cdot \omega
    \label{eq:ch7_phase_product}
\end{align}
Taking the imaginary part:
\begin{equation}
    J = |V_{us}| |V_{cb}| |V_{ub}| |V_{cs}| \cdot \sin(2\pi/3)
    \label{eq:ch7_J_formula}
\end{equation}

\textbf{Numerical result:}
\begin{align}
    J_{\text{pred}} &= 0.225 \times 0.042 \times 0.0037 \times 0.97 \times \sin(120°) \notag \\
    &= 2.93 \times 10^{-5}
    \label{eq:ch7_J_pred}
\end{align}

\textbf{Comparison:} $J_{\text{PDG}} = 3.08 \times 10^{-5}$ --- agreement within \textbf{6\%}!
\end{tcolorbox}

\paragraph{Interpretation.}
The $\mathbb{Z}_3$ structure, required independently for the generation count
(Chapter~\ref{ch:three_generations}), automatically produces CP violation of
the correct magnitude. This is a \emph{structural} result: no parameters were
adjusted to match $J$.

\subsubsection{Track B Results}

\begin{table}[ht]
\centering
\caption{Track B: CP mechanisms with one calibrated parameter}
\label{tab:ch7_trackB}
\begin{tabular}{llccl}
\toprule
\textbf{Option} & \textbf{Verdict} & \textbf{$J$ pred} & \textbf{$\delta$ pred} & \textbf{Reason} \\
\midrule
O1: Complex z-shift & RED & 0 & (cal) & Phase still removable \\
\textbf{O2: Two-path} & \textbf{YELLOW} & $3.1 \times 10^{-5}$ & 67° & Predicts $\delta$ from $|\bar\rho{-}i\bar\eta|$ \\
O3: Boundary phase & RED & 0 & N/A & All phases absorbable \\
O4: $\mathbb{Z}_3$ phases & YELLOW & $2.9 \times 10^{-5}$ & 120° & (Already works in Track A) \\
O5: Holonomy/Berry & YELLOW & $2.9 \times 10^{-5}$ & 120° & (Already works in Track A) \\
O6: Mediator mixing & RED & (cal) & (cal) & No independent prediction \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Track B insight: Option 2 (two-path interference).}
If we postulate that the $u \to b$ transition occurs via two interfering paths
with equal amplitude and calibrate the relative phase to match $|\bar\rho - i\bar\eta|$:
\begin{equation}
    |V_{ub}| = |V_{ub}|_{\text{overlap}} \cdot \cos(\phi/2)
    \implies \phi \approx 134°
    \label{eq:ch7_twopath_cal}
\end{equation}
This predicts $\delta = \phi/2 \approx 67°$, compared to PDG $\delta \approx 65°$
--- agreement within 3\%.

However, this requires postulating a second path, which is not present in the
current geometric framework.

\subsubsection{The $\delta$ Discrepancy}

Both Track~A and Track~B show a systematic issue with the CP phase $\delta$:

\begin{itemize}[nosep]
    \item Track A (O4): predicts $\delta = 120°$, PDG gives $\delta \approx 65°$
    \item Track B (O2): predicts $\delta \approx 67°$ (from calibration)
\end{itemize}

The $\mathbb{Z}_3$ mechanism gives the correct \emph{magnitude} of $J$ but the
wrong \emph{phase angle}. The factor-of-2 discrepancy in $\delta$ suggests that
the simple $\omega = e^{2\pi i/3}$ structure is an approximation. Possible
refinements:

\begin{enumerate}
    \item $\mathbb{Z}_6 = \mathbb{Z}_2 \times \mathbb{Z}_3$ selection: the $\mathbb{Z}_2$
          factor could halve certain phase contributions
    \item Non-uniform $\mathbb{Z}_3$ charges: different charge assignments for
          up vs.\ down sectors
    \item Higher-order corrections from the overlap model
\end{enumerate}

These remain (open).

\subsubsection{Attempt 3 Summary}

\begin{tcolorbox}[colback=red!10, colframe=red!60!black,
    title=\textbf{Attempt 3 Summary: FALSIFIED by Phase Cancellation Theorem}]
\begin{description}[nosep, font=\normalfont\bfseries]
    \item[Historical exploration (O4/O5):]
        An ad-hoc phase assignment gave $J = 2.9 \times 10^{-5}$, but
        \textbf{this assignment is inconsistent with any $\mathbb{Z}_3$ charge structure}.

    \item[Phase Cancellation (Attempt~4):]
        The Phase Cancellation Theorem proves that for \emph{any} valid $\mathbb{Z}_3$
        charge assignment, the Jarlskog phase sum vanishes: $J = 0$.

    \item[Why the numerical match was misleading:]
        The specific phases $(\omega^0, \omega^{-2}, \omega^{-2}, \omega^{-1})$
        used here cannot arise from charge differences $q_i^{(u)} - q_j^{(d)}$.

    \item[Resolution:]
        Attempt~4 identifies the $\mathbb{Z}_2$ factor in $\mathbb{Z}_6$ as the
        correct mechanism, yielding $\delta = 60°$, $J \simeq 2.9 \times 10^{-5}$.
\end{description}

\textbf{Status:} FALSIFIED. See Attempt~4 for the corrected analysis.
\end{tcolorbox}

\subsubsection{Updated Stoplight}

\begin{table}[ht]
\centering
\caption{Chapter 7 updated verdict after Attempt 3}
\label{tab:ch7_verdict_updated}
\begin{tabular}{lccl}
\toprule
\textbf{Claim} & \textbf{Before} & \textbf{After} & \textbf{Tag} \\
\midrule
Magnitude hierarchy $\lambda$, $\lambda^2$, $\lambda^3$ & GREEN & GREEN & \tagDc{} \\
Structure $|V_{ub}| \sim A\lambda^3$ & GREEN & GREEN & \tagDc{} \\
Prefactor $2.5\times = 1/|\bar\rho{-}i\bar\eta|$ & YELLOW & YELLOW & \tagI{} \\
Exponential profile (vs.\ Gaussian) & GREEN & GREEN & \tagDc{} \\
\midrule
\textbf{Jarlskog $J$ prediction} & \textcolor{red}{RED} & \textcolor{YellowOrange}{\textbf{YELLOW}} & \tagDc{} \\
\textbf{CP mechanism identified} & \textcolor{red}{RED} & \textcolor{YellowOrange}{\textbf{YELLOW}} & \tagDc{} \\
CP phase $\delta$ & RED & YELLOW & \tagI{} \\
$(\bar\rho, \bar\eta)$ derivation & RED & RED & (open) \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Progress summary.}
Before Attempt~3, CP violation was entirely (open)---no mechanism was identified.
After Attempt~3, the $\mathbb{Z}_3$ discrete structure provides a geometric origin
for CP violation that predicts $J$ to 6\% accuracy without calibration. The
detailed phase $\delta$ remains partially resolved ($\times 2$ discrepancy).

\subsubsection{Open Problems}

\begin{tcolorbox}[colback=gray!5, colframe=gray!50!black,
    title=\textbf{Remaining Open Problems for CP Sector}]
\begin{enumerate}[nosep]
    \item \textbf{$\delta$ prediction:}
          Z3 gives $120°$, PDG gives $65°$. Need mechanism to reduce
          effective phase by factor $\sim 2$ (Z6 structure? Z2 selection?).

    \item \textbf{$|\bar\rho - i\bar\eta|$ derivation:}
          Currently uses PDG magnitudes for $|V_{ub}|$. Need to derive the
          magnitude reduction factor from Z3 phases or interference.

    \item \textbf{Wolfenstein connection:}
          Map between Z3 phases and Wolfenstein parameters $(\bar\rho, \bar\eta)$.
          Z3 gives $\arg(\omega) = 120°$, but PDG $\arctan(\bar\eta/\bar\rho) \approx 65°$.

    \item \textbf{Z6 refinement:}
          Full $\mathbb{Z}_6 = \mathbb{Z}_2 \times \mathbb{Z}_3$ analysis may
          provide half-angle corrections or selection rules.
\end{enumerate}
\end{tcolorbox}

\subsubsection{Final Verdict}

\begin{tcolorbox}[colback=red!10, colframe=red!60!black,
    title=\textbf{Verdict: FALSIFIED --- Superseded by Attempt 4}]
\textbf{This attempt is FALSIFIED.} The Phase Cancellation Theorem (Attempt~4)
proves that pure $\mathbb{Z}_3$ charge assignments yield $J = 0$.

The numerical match reported here ($J \simeq 2.93 \times 10^{-5}$) arose from
an ad-hoc phase assignment that does not correspond to any valid $\mathbb{Z}_3$
charge configuration. While the exploration was instructive, it cannot be cited
as evidence for $\mathbb{Z}_3$-based CP violation.

\textbf{Corrected analysis (Attempt~4):}
\begin{itemize}[nosep]
    \item The $\mathbb{Z}_2$ factor in $\mathbb{Z}_6 = \mathbb{Z}_2 \times \mathbb{Z}_3$
          provides the sign selection mechanism that enables $J \neq 0$.
    \item Attempt~4 predicts $\delta = 60°$ (5° from PDG 65°) with
          $J \simeq 2.9 \times 10^{-5}$ (5\% from PDG).
    \item The $\mathbb{Z}_2$ parity assignment is identified but not fully derived.
\end{itemize}

\textbf{Bottom line:} Attempt~3 is historical exploration only.
See Section~\ref{sec:ch7_attempt4} for the active CP analysis.
\end{tcolorbox}
