% ═══════════════════════════════════════════════════════════════════════════════
% APPENDIX: DERIVATION LIBRARY
% All technical derivations used in Book 2, collected for reference.
% Updated 2026-01-29: All standalone derivations now included via .include.tex
% ═══════════════════════════════════════════════════════════════════════════════

\chapter*{Appendix: Derivation Library}
\addcontentsline{toc}{chapter}{Appendix: Derivation Library}

All technical derivations developed during the construction of Book 2 are collected
in this appendix. These derivations support the main narrative but are too detailed
for inline presentation. Each section indicates its epistemic status using standard
EDC tags.

\textbf{Organization:}
\begin{itemize}
    \item \textbf{Lemmas} --- Mathematical results reused across multiple derivations
    \item \textbf{Z$_N$ / k-Channel Derivations} --- Discrete symmetry and averaging mechanisms
    \item \textbf{BVP Physical Parameters} --- Derivations supporting $\delta$, $d_{LR}$, $f_w$
    \item \textbf{G$_F$ Non-Circular Chain} --- Framework for Fermi constant derivation
    \item \textbf{Summary Boxes} --- Quick-reference boxes for key results
\end{itemize}

% ═══════════════════════════════════════════════════════════════════════════════
% PART 1: LEMMAS
% ═══════════════════════════════════════════════════════════════════════════════
\section*{Core Lemmas}
\addcontentsline{toc}{section}{Core Lemmas}

% --- Projection-Reduction Lemma ---
\subsection*{Projection-Reduction Lemma}
\textbf{Status:} \tagDer{} \\
\textbf{Purpose:} Core lemma for mode overlap integrals in $G_F$ derivation.
Establishes Case (A) vs Case (B) for chirality structure.

\input{\EDCPAPERS/_shared/lemmas/projection_reduction_lemma}

% --- Z_6 Discrete Averaging Lemma ---
\subsection*{Z$_6$ Discrete Averaging Lemma}
\textbf{Status:} \tagDer{} \\
\textbf{Purpose:} Shows $R = 1 + a/c$ for discrete averaging over Z$_6$ anchors.
Specialization of general Z$_N$ result to hexagonal lattice.

\input{\EDCPAPERS/_shared/lemmas/z6_discrete_averaging_lemma.include}

% --- Z_N Discrete Averaging Lemma (General) ---
\subsection*{Z$_N$ Discrete Averaging Lemma (General)}
\textbf{Status:} \tagDer{} \\
\textbf{Purpose:} General result: $k(N) = 1 + 1/N$ for Z$_N$ symmetric averaging.
Foundation for the k-channel correction mechanism.

\input{\EDCPAPERS/_shared/lemmas/zn_discrete_averaging_lemma.include}

% ═══════════════════════════════════════════════════════════════════════════════
% PART 2: Z_N / K-CHANNEL DERIVATIONS
% ═══════════════════════════════════════════════════════════════════════════════
\section*{Z$_N$ / k-Channel Derivations}
\addcontentsline{toc}{section}{Z$_N$ / k-Channel Derivations}

% --- Israel Junction Z_N Fixed Points ---
\subsection*{Israel Junction Conditions at Z$_N$ Fixed Points}
\textbf{Status:} \tagDer{} \\
\textbf{Purpose:} Shows Israel junction conditions give Z$_N$ fixed points as
``identical anchors'' in the discrete averaging framework.

\input{\EDCPAPERS/_shared/derivations/israel_zn_fixed_points_anchors.include}

% --- Z_N Toy Functional from 5D Action ---
\subsection*{5D Action to Toy Functional Mapping}
\textbf{Status:} \tagDc{} \\
\textbf{Purpose:} Reduces 5D brane action to 1D toy functional with
$T = \sigma/R$, $\lambda = \kappa_5^2 \tau_n$ identification.

\input{\EDCPAPERS/_shared/derivations/zn_toy_functional_from_5d_action.include}

% --- Z_N Anisotropy Normalization ---
\subsection*{Z$_N$ Anisotropy Normalization from Energy Minimization}
\textbf{Status:} \tagDer{} \\
\textbf{Purpose:} Derives $a/c = 1/N$ from gradient energy minimization.
Key result: $a_1 \propto 1/N$ for $N$ identical anchors.

\input{\EDCPAPERS/_shared/derivations/zn_anisotropy_normalization_from_action.include}

% --- Z_N Ring Delta-Pinning Modes ---
\subsection*{Z$_N$ Ring Delta-Pinning Mode Structure}
\textbf{Status:} \tagDer{} \\
\textbf{Purpose:} BVP analysis showing $\cos(N\theta)$ is unique leading
anisotropic mode. Selection lemma: only $m = kN$ modes couple to Z$_N$ anchors.

\input{\EDCPAPERS/_shared/derivations/zn_ring_delta_pinning_modes.include}

% --- Z_N Mode Selection Nonlinear W ---
\subsection*{Z$_N$ Mode Selection under Nonlinear $W(u)$}
\textbf{Status:} \tagDer{} \\
\textbf{Purpose:} Robustness theorem: mode index $m = N$ stable under any
$C^2$ potential with stable minimum. Nonlinearities generate harmonics but
don't change mode selection.

\input{\EDCPAPERS/_shared/derivations/zn_mode_selection_nonlinear_W.include}

% --- Z_N Strong Pinning Regimes ---
\subsection*{Z$_N$ Strong Pinning Regime Analysis}
\textbf{Status:} \tagDer{} \\
\textbf{Purpose:} Shows selection lemma is geometric --- mode index $m = N$
stable at any pinning strength $\rho$. Mode shape changes (cosine $\to$ cusp)
but index doesn't.

\input{\EDCPAPERS/_shared/derivations/zn_strong_pinning_regimes.include}

% --- Z_N Symmetry Breaking One Defect ---
\subsection*{Z$_N$ Symmetry Breaking: One-Defect Robustness}
\textbf{Status:} \tagDer{} \\
\textbf{Purpose:} Perturbation theory for defect: overlap loss scales as
$O(\epsilon^2)$. k-channel robust to $\sim$10\% defect levels.

\input{\EDCPAPERS/_shared/derivations/zn_symmetry_breaking_one_defect.include}

% ═══════════════════════════════════════════════════════════════════════════════
% PART 3: BVP PHYSICAL PARAMETERS
% ═══════════════════════════════════════════════════════════════════════════════
\section*{BVP Physical Parameters}
\addcontentsline{toc}{section}{BVP Physical Parameters}

% --- Delta from 5D Action ---
\subsection*{Brane Thickness $\delta$ from 5D Action}
\textbf{Status:} \tagDc{} \\
\textbf{Purpose:} Derives $\delta = \hbar/(2m_p c) = 0.105$ fm from 5D action
mode equation. Energy minimization gives E$_0 \sim 1/\delta$; matching to proton
mass fixes scale.

\input{\EDCPAPERS/_shared/derivations/delta_from_5d_action_proton_scale.include}

% --- d_LR from Chiral Localization ---
\subsection*{L-R Separation from Chiral Localization}
\textbf{Status:} \tagDc{}/\tagCal{} \\
\textbf{Purpose:} Analyzes L-R center separation from 5D Dirac equation.
Single domain wall gives $d_{LR} = 0$; non-zero requires two walls or junction.
$G_F$ gate constrains $d_{LR}/\delta \in [5,10]$.

\input{\EDCPAPERS/_shared/derivations/dlr_from_chiral_localization.include}

% --- f_w from Stability and Spectrum ---
\subsection*{Fermion Width from Stability and Spectrum}
\textbf{Status:} \tagDc{}/\tagCal{} \\
\textbf{Purpose:} Derives fermion width window $f_w \in [0.5, 1.2]$ from
normalizability, localization, and variational constraints. Specific value
$f_w = 0.8$ calibrated from $I_4$ gate.

\input{\EDCPAPERS/_shared/derivations/fw_from_stability_and_spectrum.include}

% ═══════════════════════════════════════════════════════════════════════════════
% PART 4: G_F NON-CIRCULAR CHAIN
% ═══════════════════════════════════════════════════════════════════════════════
\section*{G$_F$ Non-Circular Chain}
\addcontentsline{toc}{section}{G$_F$ Non-Circular Chain}

% --- G_F Noncircular Chain Framework ---
\subsection*{Non-Circular $G_F$ Chain Framework}
\textbf{Status:} \tagDer{} (framework) \\
\textbf{Purpose:} Establishes forward chain for $G_F$ without using $v$
(Higgs VEV) as input. Formula: $X_{EDC} = C \cdot (g_5^2 \cdot I_4 \cdot m_e^2) / M_{eff}^2$.

\input{\EDCPAPERS/_shared/derivations/gf_noncircular_chain_framework.include}

% --- G_F Potential Shapes from 5D ---
\subsection*{BVP Potential Shapes from 5D Action Reduction}
\textbf{Status:} \tagDc{} \\
\textbf{Purpose:} Derives allowed $V(\chi)$ shapes from 5D bulk physics.
Constrains BVP background potential.

\input{\EDCPAPERS/_shared/derivations/gf_potential_shapes_from_5d.include}

% --- Prefactor A from Fluctuations ---
\subsection*{Prefactor $A$ from Semiclassical Fluctuations}
\textbf{Status:} \tagDer{} (within 1D effective model) \\
\textbf{Purpose:} Derives $A = \pi \times (\omega_0/\omega_B) / \sqrt{L_0/\delta} \approx 0.84$
connecting prefactor to brane fluctuation amplitude.

\input{\EDCPAPERS/_shared/derivations/prefactor_A_from_fluctuations.include}

% ═══════════════════════════════════════════════════════════════════════════════
% PART 5: SUMMARY BOXES
% ═══════════════════════════════════════════════════════════════════════════════
\section*{Summary Boxes}
\addcontentsline{toc}{section}{Summary Boxes}

These boxes provide quick-reference summaries of key derivations, suitable
for inline use or independent review.

\subsection*{$\delta$ from 5D Action --- Summary}
\input{\EDCPAPERS/_shared/boxes/delta_from_5d_action_box}

\subsection*{$d_{LR}$ from Chiral Localization --- Summary}
\input{\EDCPAPERS/_shared/boxes/dlr_chiral_localization_box}

\subsection*{$G_F$ BVP Pipeline --- Summary}
\input{\EDCPAPERS/_shared/boxes/gf_bvp_pipeline_box}

\subsection*{$G_F$ BVP Tuning Decomposition --- Summary}
\input{\EDCPAPERS/_shared/boxes/gf_bvp_tuning_box}

\subsection*{$N_{cell}$ Renormalization --- Summary}
\input{\EDCPAPERS/_shared/boxes/ncell_renorm_box}

\subsection*{Prefactor $A$ --- Summary}
\input{\EDCPAPERS/_shared/boxes/prefactor_A_box}

\subsection*{Z$_N$ k-Channel Robustness --- Summary}
\input{\EDCPAPERS/_shared/boxes/zn_kchannel_robustness_box}

% End of Derivation Library
