% ============================================================================
% PAPER 3 JOURNAL VERSION: Neutron Lifetime from 5D Membrane Cosmology
% ============================================================================
% BUILD: xelatex main_journal.tex && xelatex main_journal.tex
% BRANCH: restructure/paper3-journal-30p
% PURPOSE: Journal-length (~25-35 pages) with companion references
% ============================================================================
\documentclass[11pt,a4paper]{article}

% ============================================================================
% PATH B SWITCH (inherit from full paper)
% ============================================================================
\newif\ifpathB
\pathBtrue  % PATH B ENABLED

% ============================================================================
% PACKAGES
% ============================================================================
\usepackage{fontspec}

% FONTS (TeX Gyre Termes = Times-like with full OpenType support)
\IfFontExistsTF{TeX Gyre Termes}{%
  \setmainfont{TeX Gyre Termes}
  \setsansfont{TeX Gyre Heros}
}{%
  \setmainfont{Times New Roman}[Ligatures=TeX]
  \setsansfont{Helvetica}
}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{physics}
\usepackage{graphicx}
\usepackage{float}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{xcolor}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue,hypertexnames=false]{hyperref}
\usepackage[margin=2.5cm]{geometry}
\usepackage{fancyhdr}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning}
\usepackage{mathtools}
\usepackage{listings}
\usepackage{tcolorbox}
\tcbuselibrary{breakable,skins}
\usepackage{caption}

% ============================================================================
% GLOBAL TCOLORBOX STYLE
% ============================================================================
\tcbset{
    pathBbox/.style={
        width=\linewidth,
        breakable,
        enhanced,
        left=3mm,
        right=3mm,
        top=2mm,
        bottom=2mm,
        boxrule=0.5pt,
        fonttitle=\bfseries\small
    }
}

% PDF metadata
\hypersetup{
  pdftitle={Neutron Lifetime from 5D Membrane Cosmology: NJSR Edition (Journal Version)},
  pdfauthor={Igor Grcman},
  pdfsubject={EDC neutron decay model - Journal version},
  pdfkeywords={neutron lifetime, 5D membrane, EDC, brane soliton, WKB prefactor, golden ratio},
  pdfcreator={XeLaTeX},
  pdfproducer={hyperref}
}

% Suppress hyperref warnings for math in PDF bookmarks
\pdfstringdefDisableCommands{%
  \def\pi{pi}%
  \def\varepsilon{epsilon}%
  \def\epsilon{epsilon}%
  \def\Delta{Delta}%
  \def\Gamma{Gamma}%
  \def\sigma{sigma}%
  \def\hbar{hbar}%
  \def\tau{tau}%
  \def\to{->}%
  \def\alpha{alpha}%
  \def\beta{beta}%
  \def\kappa{kappa}%
  \def\lambda{lambda}%
  \def\mu{mu}%
  \def\nu{nu}%
  \def\xi{xi}%
  \def\rho{rho}%
  \def\Omega{Omega}%
  \def\delta{delta}%
  \def\gamma{gamma}%
  \def\eta{eta}%
  \def\theta{theta}%
  \def\phi{phi}%
  \def\psi{psi}%
  \def\omega{omega}%
  \def\Lambda{Lambda}%
  \def\Sigma{Sigma}%
  \def\Phi{Phi}%
  \def\Psi{Psi}%
  \def\approx{~}%
  \def\sim{~}%
  \def\equiv{=}%
  \def\neq{!=}%
  \def\leq{<=}%
  \def\geq{>=}%
  \def\cdot{.}%
  \def\times{x}%
  \def\pm{+-}%
  \def\infty{inf}%
  \def\partial{d}%
  \def\nabla{grad}%
  \def\sqrt#1{sqrt(#1)}%
  \def\frac#1#2{#1/#2}%
  \def\tfrac#1#2{#1/#2}%
  \def\mathrm#1{#1}%
  \def\text#1{#1}%
  \def\textbf#1{#1}%
  \def\mathbf#1{#1}%
  \def\boldsymbol#1{#1}%
  \def\rm{}%
  \def\_{}%
}

% ============================================================================
% EPISTEMIC MARKERS
% ============================================================================
\newcommand{\Dc}{\textcolor{green!60!black}{\textbf{[Dc]}}}
\newcommand{\Pp}{\textcolor{orange}{\textbf{[P]}}}
\newcommand{\Dd}{\textcolor{blue}{\textbf{[D]}}}
\newcommand{\Mm}{\textcolor{purple}{\textbf{[M]}}}
\newcommand{\Ii}{\textcolor{gray}{\textbf{[I]}}}
\newcommand{\BL}{\textcolor{brown}{\textbf{[BL]}}}
\newcommand{\Cal}{\textcolor{cyan!60!black}{\textbf{[Cal]}}}
\newcommand{\Open}{\textcolor{red}{\textbf{[OPEN]}}}

% ============================================================================
% SM BASELINE TOGGLE
% ============================================================================
\newif\ifSMbaseline
\SMbaselinetrue

\newenvironment{SMbaselineblock}{%
    \ifSMbaseline
    \begin{tcolorbox}[colback=brown!5!white, colframe=brown!50!black,
        title={\small\BL\ Standard Model Baseline (for comparison only)},
        fonttitle=\bfseries\small, breakable]
}{%
    \end{tcolorbox}
    \fi
}

\newcommand{\SMterm}[1]{\ifSMbaseline\textcolor{brown}{#1}\BL\else[SM]\fi}

% ============================================================================
% THEOREM ENVIRONMENTS
% ============================================================================
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{postulate}{Postulate}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% ============================================================================
% CUSTOM COMMANDS
% ============================================================================
\newcommand{\VolB}{\mathrm{Vol}(B^3)}
\newcommand{\AreaS}{\mathrm{Area}(S^3)}
\newcommand{\Plenum}{\mathcal{P}}
\newcommand{\Membrane}{\Sigma^3}
\newcommand{\Bulk}{\mathcal{M}^5}
\newcommand{\Rdet}{R_{\mathrm{det}}}
\newcommand{\phitail}{\varphi_{\mathrm{tail}}}

% Framework naming
\newcommand{\FrameworkNameLong}{Neutron Junction-Slip Reduction}
\newcommand{\FrameworkNameShort}{NJSR}
\newcommand{\FrameworkNameEdition}{\FrameworkNameShort\ Edition}
\newcommand{\FrameworkNameFramework}{\FrameworkNameLong\ (\FrameworkNameShort) framework}

% Canonical notation: bulk-brane exchange current (Framework v2.0)
\newcommand{\Jbb}[1]{J^{#1}_{\mathrm{bulk}\to\mathrm{brane}}}

% ============================================================================
% DOCUMENT INFO
% ============================================================================

\title{%
\textbf{Neutron Lifetime from 5D Membrane Cosmology} \\[0.3cm]
\Large{WKB Tunneling, Brane-Soliton Structure, and the \FrameworkNameShort\ Framework} \\[0.5cm]
\large{\textcolor{purple}{Paper 3 --- \FrameworkNameEdition\ (Journal Version)}}
}

\author{
Igor Gr\v{c}man\footnote{Corresponding author. Email: igor.grcman@gmail.com}
\\[0.2cm]
\small{with computational assistance from AI}
\\[0.5cm]
\small{Elastic Diffusive Cosmology Research}
}

\date{January 2026}

% ============================================================================
\begin{document}
% ============================================================================

\maketitle

% Epigraphs
\begin{quote}\itshape
``Jedenfalls bin ich \"uberzeugt, da{\ss} der Alte nicht w\"urfelt.''\\
\hfill--- Albert Einstein, letter to Max Born (4 Dec 1926)~\cite{EinsteinBorn1926}\\[0.35em]
No --- God plays with compass and triangles.\\
\hfill--- Igor Gr\v{c}man (author paraphrase)
\end{quote}

\begin{center}
\small DOI: \href{\PaperDOIURL}{\PaperDOI}\\[0.2cm]
\small Repository: \url{https://github.com/igorgrcman/elastic-diffusive-cosmology}\\
\small\textit{(Public artifacts in \texttt{edc\_papers} folder)}
\end{center}

\vspace{0.3cm}
\begin{center}
\small\textbf{Related Documents:}\\[0.15cm]
\footnotesize
\textbf{Companions:}\\
\end{center}

% ============================================================================
% ABSTRACT
% ============================================================================
\begin{abstract}
We present the \textbf{\FrameworkNameFramework} for computing particle lifetimes from
5D Elastic Diffusive Cosmology (EDC), with the neutron as the primary test case.
The framework provides:
\begin{enumerate}
    \item A WKB tunneling calculation with explicit prefactor $A_0 = (\omega_{\rm well}/2\pi) \cdot \Rdet \cdot C_{\rm zero}$
    \item Brane-bound soliton structure for the electron ($Q = -1$ defect) with
          golden-ratio tail exponent $\phitail = (1+\sqrt{5})/2 \approx 1.618$
    \item Determinant ratio $\Rdet = 0.63 \pm 0.10$ \Dc\ from Gel'fand--Yaglom analysis (method-spread systematic)
    \item Decay output classification (brane-bound vs bulk-escape) without Standard Model dynamics
\end{enumerate}

\medskip
\noindent\textbf{Calibration statement:} We calibrate a single barrier-height scale $V_B$ to the measured
free-neutron lifetime $\tau_n = 878.4 \pm 0.5$\,s \BL. All other outputs---the functional forms of
$V(q)$ and $M(q)$ under stated ansätze \Dc, the golden-ratio tail exponent $\phitail$ \Dc, the
determinant ratio $\Rdet$ \Dc\ (method-spread systematic), and 10/10 verification gates---are
\emph{not fitted} but follow from the reduction pipeline with $V_B$ held fixed.

\medskip
\noindent\textbf{This is a journal-length version.} For full forensic derivation chains, verification gates,
and worked calculations, see the Companion Papers referenced in Section~\ref{sec:companions}.

\medskip
\noindent\textbf{Guardrail:} \FrameworkNameShort\ uses \textbf{5D EDC language only}---no V--A structure,
no W boson, no CKM matrix. Standard Model observables serve as calibration benchmarks \BL,
not as inputs to the derivation.
\end{abstract}

\vspace{0.5cm}
\noindent\textbf{Keywords:} neutron lifetime, 5D membrane cosmology, WKB tunneling,
brane soliton, golden ratio, EDC, determinant ratio

\tableofcontents
\newpage

% ============================================================================
% PHANTOM LABELS for appendix references in main_body.tex
% These resolve refs that point to full-paper appendices → journal companions
% ============================================================================
% These labels must be defined BEFORE main_body is input so refs resolve correctly
\phantomsection\label{app:neutron_lifetime}      % Referenced in main_body → See Companion B
\phantomsection\label{app:5d_reduction}          % Referenced in main_body → See Companion C
\phantomsection\label{app:worked_derivation}     % Referenced in main_body → See Companion A
\phantomsection\label{app:sign_derivation}       % Referenced in main_body → See summary appendix
% Phantom table label for gates table (main_body references Table~\ref{tab:gates})
\phantomsection\label{tab:gates}

% ============================================================================
% MAIN BODY (shared with full paper)
% ============================================================================
\input{body_shared/main_body}

% ============================================================================
% COMPANION PAPERS SECTION (NEW FOR JOURNAL VERSION)
% Note: Roadmap section is included via main_body.tex
% ============================================================================
\section{Companion Papers and Full Derivations}
\label{sec:companions}

This journal-length paper presents the core narrative and key results. For full derivation
chains, verification gates, and worked calculations, see the following companion documents:

\begin{table}[H]
\centering
\small
\begin{tabular}{clp{7cm}}
\toprule
\textbf{Document} & \textbf{Title} & \textbf{Content} \\
\midrule
Framework v2.0 & EDC Mathematical Framework &
    Complete 5D theory: mass formulas, $\alpha$, Y-junction structure, SU(3) color.
Companion A & Effective Lagrangian $L_{\rm eff}$ from 5D Action &
    Complete derivation of $L_{\rm eff}(q,\dot{q}) = \frac{1}{2}M(q)\dot{q}^2 - V(q)$
    via Israel junction conditions, with $M(q)$ and $V(q)$ explicit forms.
Companion B & WKB Prefactor and Lifetime Calculation &
    Gel'fand--Yaglom determinant ratio $R_{\rm det} = 0.63 \pm 0.10$, golden-ratio
    tail exponent $\varphi = 1.618...$, and 10/10 verification gates.
Companion C & 5D Reduction Pipeline &
    Roadmap from brane-world action to 1D mechanics: bulk geometry, embedding,
    induced metric, extrinsic curvature, and assumption ledger.
Companion D & Selection Rules for $\beta^-$ Decay &
    Topological necessity of neutral mode with $p^\xi \neq 0$; charge, winding,
    and momentum budget bookkeeping; antineutrino classification.
Companion E & Symmetry Layering and Defect Operations &
    Formal symmetry structure, process operators (Excitation, Relaxation, Merging),
    conservation ledger for $\beta^-$ decay.
\bottomrule
\end{tabular}
\caption{Companion papers providing full derivation chains. These documents contain the
\label{tab:companions}
\end{table}

\noindent\textbf{Navigation guide:}
\begin{itemize}
    \item \textbf{This paper} ($\sim$25 pages): narrative, key results, and conclusions---sufficient
          for understanding the claim and its epistemic status.
    \item \textbf{Framework v2.0} (38 pages): complete EDC mathematical framework.
    \item \textbf{Companions A--E} ($\sim$50 pages total): full forensic derivation chain---for readers
          who wish to audit every step.
    \item \textbf{Full Paper 3} (147 pages): complete document with all appendices inline, available
          as \texttt{main\_pathB.pdf} in the repository.
\end{itemize}

% ============================================================================
% TECHNICAL SUMMARY
% ============================================================================
\section{\FrameworkNameShort: Technical Summary}
\label{sec:tech_summary}

This section provides a condensed technical summary of the \FrameworkNameShort\ framework.

\subsection{Language Guardrail}

The \FrameworkNameShort\ framework uses \textbf{5D EDC language only}:
\begin{itemize}
    \item No V--A structure (replaced by junction-slip topology)
    \item No W boson propagator (replaced by collective-coordinate tunneling)
    \item No CKM matrix (replaced by geometric phase factors)
\end{itemize}
Standard Model observables serve as calibration benchmarks \BL, not as inputs.

\subsection{Module Summary}

The calculation pipeline consists of the following modules:
\begin{enumerate}
    \item \textbf{5D Action} $\to$ collective coordinate $q$ (Companion~C: \emph{5D Action Reduction Pipeline})
    \item \textbf{Dimensional reduction} $\to$ $S_{\rm eff}[q] = \int dt\, \left(\frac{1}{2}M(q)\dot{q}^2 - V(q)\right)$ (Companion~A: \emph{Effective Lagrangian})
    \item \textbf{WKB tunneling} $\to$ decay rate $\Gamma = A_0 e^{-2B}$ (Companion~B: \emph{WKB Prefactor})
    \item \textbf{Calibration} $\to$ $V_B$ fitted to $\tau_n = 878.4$\,s \Cal
\end{enumerate}

\subsection{Key Results}

\begin{enumerate}
    \item Golden-ratio tail exponent: $\phitail = (1+\sqrt{5})/2$ \Dc
    \item Determinant ratio: $\Rdet = 0.63 \pm 0.10$ \Dc
    \item Verification gates: 10/10 passed \Dc
    \item Single calibration parameter: $V_B$ \Cal
\end{enumerate}

\subsection{Narrative Arc: Symmetry Through Saturation}
\label{subsec:narrative_arc}

The EDC program unfolds through three conceptual stages, each with distinct
epistemic status:

\paragraph{Act I: $(e, p)$ Topology Distinction \Dc.}
The electron and proton represent fundamentally different membrane defect classes.
The electron is a surface defect (simple vortex, $B^3$ configuration space, winding $W = -1$);
the proton is a volume defect (Y-junction, $S^3 \times S^3 \times S^3$ configuration space).
This topological distinction yields the mass ratio $m_p/m_e = 6\pi^5$ (0.01\% agreement)
and explains why the electron is brane-confined while baryons have flux-tube extensions.

\paragraph{Act II: Muon as Higher-Overlap Generational Mode \Pp.}
The muon is postulated to be a second-generation leptonic defect characterized by
increased ``overlap'' with baryonic internal geometry. In this picture, the muon
samples a larger portion of the Y-junction configuration space than the electron,
yielding higher effective mass. The mass relation $m_\mu/m_e = \frac{3}{2}(1 + \alpha^{-1})$
(0.14\% agreement) is consistent with this overlap interpretation, but the mechanism
remains \Pp---a conjecture awaiting derivation from the 5D action.

\paragraph{Act III: Tau and the Saturation Hypothesis \Pp.}
The tau represents maximal leptonic overlap with baryonic geometry before topological
instability sets in. The \textbf{Saturation Hypothesis} \Pp\ proposes that no further
stable lepton generation exists because the tau already saturates the available
overlap regime. This is explicitly a conjecture: the ``saturation'' criterion has not
been derived from the action, and the absence of a fourth generation is imposed, not proven.

\medskip
\noindent\textbf{Status summary:} Act~I is derived (\Dc); Acts~II and III are
postulated (\Pp) and require future derivation to upgrade their epistemic standing.

\subsection{Strong-Sector Vocabulary}
\label{subsec:strong_sector}

The Y-junction model provides natural mappings for strong-sector concepts
(see~\cite{EDC_Paper2} for detailed derivations):
\begin{itemize}
    \item \textbf{Color (SU(3)):} The 8 gluon modes emerge from Y-junction arm
          permutation algebra \Dc---the junction has $\mathbb{Z}_6$ symmetry generating
          8 independent operations matching the 8 generators of SU(3).
    \item \textbf{Quarks:} Leg endpoints of the Y-junction; confined by infinite
          string energy \Pp.
    \item \textbf{Spin, helicity, chirality:} Conserved quantities tracked via the
          conservation ledger \BL. These are QFT inputs, not re-derived within EDC;
          ledger closure ensures consistency across decay channels.
\end{itemize}

\subsection{Adopted Values}

\begin{table}[H]
\centering
\small
\begin{tabular}{lcll}
\toprule
\textbf{Quantity} & \textbf{Value} & \textbf{Status} & \textbf{Source} \\
\midrule
$\tau_n$ & $878.4 \pm 0.5$\,s & \BL & PDG 2024 \\
$\phitail$ & $1.618...$ & \Dc & Asymptotic ODE analysis \\
$\Rdet$ & $0.63 \pm 0.10$ & \Dc & Gel'fand--Yaglom \\
$V_B$ & (calibrated) & \Cal & Fitted to $\tau_n$ \\
\bottomrule
\end{tabular}
\caption{Key quantities and their epistemic status in the \FrameworkNameShort\ framework.}
\label{tab:adopted}
\end{table}

% ============================================================================
% SUMMARY APPENDICES (short, essential only)
% ============================================================================
\appendix

\section{Neutron Instability and Beta Decay (Summary)}
\label{app:neutron}

The neutron is modeled as an asymmetric Y-junction in the 5D membrane framework.
Unlike the proton (symmetric Y-junction, stable), the neutron has stored
configurational energy $\Delta E_{\rm config} > 0$ that can be released through
a junction-slip transition.

The decay $n \to p + e^- + \bar{\nu}_e$ is interpreted as a collective-coordinate
tunneling event where the asymmetric junction relaxes to the symmetric configuration
while emitting brane-bound (electron) and bulk-escape (antineutrino) defects.


\section{Matter-Antimatter Asymmetry (Summary)}
\label{app:matter_antimatter}

The INFLOW/OUTFLOW framework proposes that cosmological particle creation draws
energy from the 5D Plenum, eliminating the need for antimatter as a conservation
partner. Under assumptions A1--A6:
\begin{align}
\delta S_{\rm INFLOW} &< 0 \quad \text{(energetically favored)} \\
\delta S_{\rm OUTFLOW} &> 0 \quad \text{(energetically suppressed)}
\end{align}

This replaces the Sakharov conditions with Plenum properties. The sign derivation
remains \Open\ pending a first-principles calculation.

\textbf{For full derivation:} See the full Paper 3 Appendix B.

\section{Mathematical Formalization (Summary)}
\label{app:formalization}

The INFLOW/OUTFLOW formalism rests on assumptions A1--A6:
\begin{enumerate}
    \item[A1.] Plenum has positive pressure: $P_{\rm bulk} > 0$
    \item[A2.] Energy conservation in 5D: $\nabla_A T^{AB} = 0$
    \item[A3.] Membrane is a codimension-1 hypersurface
    \item[A4.] Defects are localized perturbations of the membrane
    \item[A5.] INFLOW defects have $J^\xi > 0$ (energy flows into membrane)
    \item[A6.] OUTFLOW defects have $J^\xi < 0$ (energy flows out of membrane)
\end{enumerate}

\medskip\noindent
\textbf{Canonical energy-exchange statement} (Framework v2.0, Remark~4.5):
\[
\nabla_A T^{AB}_{(5)}=0\ (A,B=0,\dots,4),\qquad
\nabla_\mu T^{\mu\nu}_{\mathrm{brane}}=-\,\Jbb{\nu}\ (\mu,\nu=0,\dots,3).
\]
Sign convention: $\Jbb{\nu}>0$ denotes bulk$\to$brane (INFLOW).

\emph{This block is quoted verbatim from Framework v2.0, Remark~4.5; Framework remains the canonical source.}

\textbf{For full derivation:} See the full Paper 3 Appendix C.

\section{Collective Coordinate $q$: Geometric Definition (Summary)}
\label{app:q_geometry}

The collective coordinate $q$ is defined as the normalized ``openness'' of the
junction defect:
\begin{equation}
q \equiv \frac{A_{\rm junction}}{A_0}
\end{equation}
where $A_{\rm junction}$ is the junction cross-sectional area and $A_0$ is a
reference scale (electron Compton wavelength squared).

The range is $q \in [0, 1]$ where:
\begin{itemize}
    \item $q = 0$: closed junction (stable proton)
    \item $q = 1$: fully open junction (decay threshold)
\end{itemize}


% ============================================================================
% ACKNOWLEDGMENTS
% ============================================================================
\section*{Acknowledgments}

This work builds on the EDC framework developed in the \emph{EDC Theory Book} v17.49~\cite{EDC_Book}
and the companion paper~\cite{EDC_Paper2}.
Terminology and the 3D/4D $\leftrightarrow$ 5D mapping follow the \emph{Framework Reference Document v2.0}~\cite{grcman2026_framework_v2}.

% ============================================================================
% BIBLIOGRAPHY
% ============================================================================
\begin{thebibliography}{99}

\bibitem{EDC_Book}
I.~Gr\v{c}man, ``Elastic Diffusive Cosmology: A 5D Membrane Framework for
Fundamental Physics,'' version 17.49, Zenodo (2026).

\bibitem{EDC_Paper2}
I.~Gr\v{c}man, ``Geometric Structure of Electron and Proton in 5D Membrane
Cosmology: Derivation of the Fine-Structure Constant,'' Zenodo (2026).

% Note: EDCPartI_Zenodo and EDCAlphaReport_Zenodo removed (duplicates of EDC_Book and EDC_Paper2)

\bibitem{EinsteinBorn1926}
A.~Einstein and M.~Born, \emph{The Born--Einstein Letters: Correspondence between Albert Einstein and Max and Hedwig Born 1916--1955}, translated by I.~Born (Macmillan, London, 1971), letter dated 4 December 1926.

\bibitem{Sakharov1967}
A.~D.~Sakharov, ``Violation of CP invariance, C asymmetry, and baryon
asymmetry of the universe,'' JETP Lett. \textbf{5}, 24 (1967).

\bibitem{PDG2024}
Particle Data Group, ``Review of Particle Physics,''
Phys. Rev. D \textbf{110}, 030001 (2024).

\bibitem{AMS02}
AMS Collaboration, ``The Alpha Magnetic Spectrometer (AMS) on the
International Space Station,'' Phys. Rep. \textbf{894}, 1 (2021).

\bibitem{FermiLAT}
Fermi-LAT Collaboration, ``Search for gamma-ray emission from
dark matter annihilation,'' Phys. Rev. D \textbf{91}, 122002 (2015).

\bibitem{FermiGamma}
Fermi-LAT Collaboration, ``Constraints on Galactic dark matter
annihilation from gamma-ray observations,'' Phys. Rev. D \textbf{103}, 083016 (2021).

% Companion paper references
\bibitem{Companion_A}
I.~Gr\v{c}man, ``Derivation of the Effective Lagrangian $L_{\rm eff}(q,\dot{q})$
from the 5D Einstein-Hilbert Action,''
Zenodo (2026). [Companion A to Paper 3]

\bibitem{Companion_B}
I.~Gr\v{c}man, ``WKB Prefactor and Neutron Lifetime Calculation
in Elastic Diffusive Cosmology,''
Zenodo (2026). [Companion B to Paper 3]

\bibitem{Companion_C}
I.~Gr\v{c}man, ``5D Action Reduction Pipeline:
From Brane-World Gravity to Effective 1D Collective Dynamics,''
Zenodo (2026). [Companion C to Paper 3]

\bibitem{Companion_D}
I.~Gr\v{c}man, ``Selection Rules for Neutron $\beta^-$ Decay
in Elastic Diffusive Cosmology,''
Zenodo (2026). [Companion D to Paper 3]

\bibitem{grcman2026_framework_v2}
I.~Gr\v{c}man, ``EDC 5D Complete Mathematical Framework: Reference Document v2.0,''
Zenodo (2026).

\end{thebibliography}

% ============================================================================
\end{document}
% ============================================================================
