\chapter{Master Tables of Fundamental Physics Problems}
\label{app:mastertables}

This appendix presents comprehensive catalogs of the problems discussed in Chapter 1, organized by domain.

\section{Quantum Mechanics Problems}

\definecolor{headerblue}{RGB}{51,102,153}
\definecolor{rowgray}{RGB}{245,245,245}

\begin{longtable}{>{\raggedright\arraybackslash}p{3cm} >{\centering\arraybackslash}p{1.2cm} >{\raggedright\arraybackslash}p{10cm}}
\caption{Quantum Mechanics Crisis Catalog} \\
\toprule
\rowcolor{headerblue}
\textcolor{white}{\textbf{Problem}} & \textcolor{white}{\textbf{Type}} & \textcolor{white}{\textbf{Description \& Status}} \\
\midrule
\endfirsthead

\toprule
\rowcolor{headerblue}
\textcolor{white}{\textbf{Problem}} & \textcolor{white}{\textbf{Type}} & \textcolor{white}{\textbf{Description \& Status}} \\
\midrule
\endhead

\midrule
\endfoot

\bottomrule
\endlastfoot

\textbf{Wavefunction Ontology} & I & 
\textbf{Crisis:} What is $\psi$? Copenhagen: "probability amplitude" --- circular definition. No physical substrate. \\
& & \textbf{Status:} All interpretations (Copenhagen, Many-Worlds, Pilot-Wave, GRW) philosophically unsatisfying. \\
& & \textbf{Required:} Physical substrate; derive Born rule from geometry. \\
& & \textbf{EDC Solution:} $\psi$ = membrane displacement field; Born rule from diffusive dynamics (Chapter 7). \\
\midrule
\rowcolor{rowgray}

\textbf{Measurement Problem} & I & 
\textbf{Crisis:} Two incompatible dynamics: Schrödinger (deterministic) vs collapse (stochastic). Von Neumann's consciousness solution. \\
& & \textbf{Status:} Schrödinger cat, Wigner friend experiments. Decoherence doesn't solve true collapse. \\
& & \textbf{Required:} Unified evolution; no ad-hoc quantum-classical cut. \\
\midrule

\textbf{Heisenberg Uncertainty} & I & 
\textbf{Crisis:} Is $\Delta x \Delta p \geq \hbar/2$ epistemic (ignorance) or ontic (fundamental)? No geometric derivation. \\
& & \textbf{Status:} Bell tests rule out local hidden variables ($>100\sigma$); 2022 Nobel Prize. \\
& & \textbf{Required:} Derive from dimensional projection; explain commutation relations geometrically. \\
& & \textbf{EDC Solution:} Ontic; emerges from vortex dynamics on elastic membrane (Chapter 7). \\
\midrule
\rowcolor{rowgray}

\textbf{Entanglement} & I & 
\textbf{Crisis:} Bell inequality violations prove non-local correlations. No mechanism. "Spooky action at a distance." \\
& & \textbf{Status:} Loophole-free tests (Hensen, Giustina, Shalm 2015). No-signaling theorem operational fix. \\
& & \textbf{Required:} Geometric connection through higher dimension; local in 5D → non-local in 3D. \\
\midrule

\textbf{Renormalization} & T & 
\textbf{Crisis:} Point particles → infinite self-energies. "Solution": subtract infinities ($\infty - \infty$). \\
& & \textbf{Status:} Feynman: "dippy process"; Dirac: "senseless mathematics." Works but philosophically bankrupt. \\
& & \textbf{Required:} Finite-size structures; UV-complete theory without divergences. \\

\end{longtable}

\section{Cosmology Problems}

\begin{longtable}{>{\raggedright\arraybackslash}p{3cm} >{\centering\arraybackslash}p{1.2cm} >{\raggedright\arraybackslash}p{10cm}}
\caption{Cosmology Crisis Catalog} \\
\toprule
\rowcolor{headerblue}
\textcolor{white}{\textbf{Problem}} & \textcolor{white}{\textbf{Type}} & \textcolor{white}{\textbf{Description \& Status}} \\
\midrule
\endfirsthead

\toprule
\rowcolor{headerblue}
\textcolor{white}{\textbf{Problem}} & \textcolor{white}{\textbf{Type}} & \textcolor{white}{\textbf{Description \& Status}} \\
\midrule
\endhead

\midrule
\endfoot

\bottomrule
\endlastfoot

\textbf{Dark Matter} & E & 
\textbf{Crisis:} Flat rotation curves require 85\% of matter to be invisible. Zero direct detection in 40 years (LUX, XENON, LZ null). \\
& & \textbf{Status:} "Epicycles of 21st century" --- parameter space shrinking, no discovery. Bullet Cluster cited as "proof" but evidence is circular. \\
& & \textbf{Required:} Emergent from modified dynamics or higher-dimensional geometry; not new particles. \\
& & \textbf{EDC Solution:} Plenum drag on galactic scales mimics dark matter halo (River Model, Chapter 9). \\
\midrule
\rowcolor{rowgray}

\textbf{Vacuum Catastrophe} & T+E & 
\textbf{Crisis:} QFT predicts $\rho_{vac}^{theory}/\rho_{vac}^{obs} \sim 10^{120}$. Worst prediction in history. \\
& & \textbf{Status:} Fine-tuning to 120 decimal places. Anthropic principle = miracle, not explanation. \\
& & \textbf{Required:} Dynamical vacuum energy; resolve catastrophe without fine-tuning. \\
& & \textbf{EDC Solution:} Membrane tension $\sigma_{eff}$ yields $\Lambda \sim \sigma_{eff} r_e / R_H^2$ --- correct order of magnitude (Chapter 12, Epilogue). \\
\midrule

\textbf{Inflation} & T & 
\textbf{Crisis:} Solves horizon/flatness but requires ad-hoc inflaton field. 30+ models, all fitting data by parameter tuning. Unfalsifiable (Popper). \\
& & \textbf{Status:} Steinhardt (co-inventor): "too flexible to be scientific." Alternative: VSL dismissed as "crazy." \\
& & \textbf{Required:} VSL or causal mechanism without new scalar field. \\
\midrule
\rowcolor{rowgray}

\textbf{Hubble Tension} & E & 
\textbf{Crisis:} CMB (Planck): $H_0 = 67.4$ km/s/Mpc vs Cepheids (SH0ES): $H_0 = 73.0$ km/s/Mpc. 5-6$\sigma$ discrepancy. \\
& & \textbf{Status:} Both teams checked exhaustively. Riess: "We need new physics." Not going away. \\
& & \textbf{Required:} Modified expansion history or breakdown of $\Lambda$CDM. \\

\end{longtable}

\section{Foundational Problems}

\begin{longtable}{>{\raggedright\arraybackslash}p{3cm} >{\centering\arraybackslash}p{1.2cm} >{\raggedright\arraybackslash}p{10cm}}
\caption{Foundational Problems Catalog} \\
\toprule
\rowcolor{headerblue}
\textcolor{white}{\textbf{Problem}} & \textcolor{white}{\textbf{Type}} & \textcolor{white}{\textbf{Description \& Status}} \\
\midrule
\endfirsthead

\toprule
\rowcolor{headerblue}
\textcolor{white}{\textbf{Problem}} & \textcolor{white}{\textbf{Type}} & \textcolor{white}{\textbf{Description \& Status}} \\
\midrule
\endhead

\midrule
\endfoot

\bottomrule
\endlastfoot

\textbf{Circular Definitions} & I & 
\textbf{Crisis:} Mass defined via force ($F=ma$), force via mass. Charge via EM field, field via charge. Operational definitions hide ontological void. \\
& & \textbf{Status:} Mach's Principle forgotten. No geometric grounding. \\
& & \textbf{Required:} Derive mass, charge, force from geometry or topology. \\
\midrule
\rowcolor{rowgray}

\textbf{Forbidden Questions} & I & 
\textbf{Crisis:} "Before Big Bang" = meaningless. "Between measurements" = meaningless. "Mechanism of entanglement" = meaningless. \\
& & \textbf{Status:} Instrumentalism ("shut up and calculate") prevents ontological inquiry. \\
& & \textbf{Required:} Answer questions mechanically, not dismiss as "meaningless." \\
& & \textbf{EDC Solution:} Wavefunction = membrane vibration; "before Big Bang" = Plenum state before scanning (Epilogue). \\
\midrule

\textbf{Fine-Tuning} & I & 
\textbf{Crisis:} $\sim$30 free parameters in SM. Why $m_e = 0.511$ MeV? Why $\alpha = 1/137.036$? Anthropic principle = non-explanation. \\
& & \textbf{Status:} All measured precisely; none derived theoretically. \\
& & \textbf{Required:} Derive constants from geometry, dynamics, or topology. \\
& & \textbf{EDC Solution:} Three-Scale Hierarchy ($\ell_P$, $R_\xi$, $r_e$) derives $\hbar$, $\alpha$, $G$ from membrane geometry (Chapters 7--8). \\

\end{longtable}

\vspace{0.5cm}
\noindent\textbf{Legend:}
\begin{itemize}[leftmargin=*,nosep]
    \item \textbf{Type I} = Interpretational/Foundational
    \item \textbf{Type T} = Theoretical Consistency
    \item \textbf{Type E} = Empirical Tension
\end{itemize}

\newpage




