\chapter*{Glossary}
\addcontentsline{toc}{chapter}{Glossary}

\begin{description}[style=nextline,leftmargin=3cm]
    \item[Acoustic Horizon] 
    The maximum distance sound waves (pressure perturbations) could have traveled in the Plenum since the Big Bang, analogous to the particle horizon in standard cosmology.
    
    \item[Baryon Acoustic Oscillations (BAO)]
    Regular fluctuations in the density of baryonic matter caused by acoustic waves in the early universe, used as a standard ruler in cosmology.
    
    \item[Bulk]
    The 5-dimensional energetic fluid (Plenum) in which the 3D membrane is embedded. Also called ``hyperspace'' or ``fifth dimension.''
    
    \item[Cosmological Principle]
    The assumption that the universe is homogeneous (same density everywhere) and isotropic (same in all directions) on large scales ($> 300$ Mly).
    
    \item[EDC (Elastic Diffusive Cosmology)]
    A theoretical framework treating the universe as a 3D tensioned membrane embedded in a 5D viscous fluid (Plenum), deriving all known physics from membrane-Plenum dynamics.
    
    \item[Flux Event]
    A discrete energy transfer event when 5D energy flux crosses from the Bulk into the 3D membrane, perceived as a ``particle'' or ``photon.''
    
    \item[Horizon Problem]
    The puzzle that causally disconnected regions of the universe (separated by more than 2° on the CMB sky) have nearly identical temperatures.
    
    \item[Membrane]
    The 3-dimensional spatial manifold we inhabit, modeled as an elastic sheet with tension $\sigma$ embedded in the 5D Bulk.
    
    \item[Plenum]
    The 5-dimensional energetic fluid filling the Bulk, characterized by viscosity $\eta_{\text{bulk}}$ and energy density $\rho_{\text{bulk}}$.
    
    \item[Renormalization]
    A mathematical procedure in quantum field theory where infinite self-energies are ``subtracted'' by adjusting bare parameters to match observed values. Feynman called it ``hocus-pocus.''
    
    \item[Tidal Charge]
    A geometric deformation of the membrane induced by Bulk pressure gradients, producing gravitational effects without requiring mass as fundamental.
    
    \item[Wave-Particle Duality]
    The apparent dual nature of quantum objects; in EDC, this emerges from the projection of continuous 5D flux structures onto the 3D membrane.
    
    \item[WIMP (Weakly Interacting Massive Particle)]
    A hypothetical dark matter candidate that interacts only via gravity and the weak nuclear force. After 40 years of searching, none have been detected.
\end{description}

